---
title: "Time Series"
subtitle: "Fahrmeier Statistik Chapter 14 'Zeitreihen"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
output:
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 12
    number_sections: yes
    toc: yes
  html_document: default
  word_document:
    toc: yes
  odt_document:
urlcolor: blue
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initialization, echo = FALSE, message = FALSE, warning = FALSE}
library(zoo)
library(tseries) # Finance data can be easily queried using the get.hist.quote
library(stlplus) # Enhanced Seasonal Decomposition of Time Series by Loess
library(tidyverse)
library(magrittr)
library(tsibble)
# for package overview see:
#                   https://cran.r-project.org/web/views/TimeSeries.html
#                   https://slides.earo.me/rstudioconf19/#16
library(fable)
# vignette("fable")
library(gridExtra)

#################################
# library(devtools)
# # devtools::load_all("animals4") # loads package incl. data files (if available)
# 
# devtools::install("D:/Wolfgang/Programs-R/statistikv8")
# devtools::load_all("D:/Wolfgang/Programs-R/statistikv8")
#                             # loads package incl. data files (if available)
# library("statistikv8")
# help("abb14_loess")
# dir("D:/Wolfgang/Programs-R/statistikv8/R/", full.names = TRUE)
#################################

theme_replace(plot.title = element_text(hjust = 0.5, face = "bold.italic", 
                                        color = "darkcyan", size = 12),
              plot.subtitle = element_text(hjust = 0.5, face = "italic", 
                                           color = "darkcyan"))

```


# Chapter 14.1 -14.5 Zeitreihen (p. 503ff)

## BMW Stock Values from file bmw.txt

Covers Abb. 14.1, 14.8 & 14.9

```{r  bmw txt, echo = FALSE, fig.width= 7, fig.height= 8}

# data from
# http://chris.userweb.mwn.de/statistikbuch/luftschad.txt

(bmw <- as_tibble(read.table("Data_Fahrmeir/bmw.txt", header = TRUE)))
# bmw$new_Date <-  # converting Date-time from character vector to class "POSIXct"
#   as.POSIXct(strptime(bmw$Date, "%Y-%m-%d"))
bmw$Date <- as.Date(bmw$Date)
bmw <- as_tsibble(bmw, index = Date)
ordnung <- 201
bmw %<>% mutate(rol_mean = 
         stats::filter(Tageskurs, filter = rep(1/ordnung, ordnung)))



bmw_2000_2015 <- ggplot(bmw, aes(x = Date, y = Tageskurs)) +
  # geom_point(col = "blue", na.rm = TRUE) +
  geom_line(aes(col = "Daily Stock Prices"), linetype = "solid", na.rm = TRUE) +
  geom_line(aes(y = rol_mean, col = "Rolling Mean"), linetype = "solid", size = 1.3,
            na.rm = TRUE) +
  geom_smooth(aes(col = "Loess Reg"), method = "loess", size = 1, na.rm = TRUE) +
  labs(x = "Date", y = "Stock Price / Euro", col = "") +
  ggtitle("Dayly BMW Stock Prices (see Fahrmeir Abb. 14.9, p.519)",
          subtitle = paste("Period of Time: ", min(bmw$Date), "-", max(bmw$Date),
                           "; Rolling Mean: ", ordnung, "days")) +
  # limits: ordering of colours
	# values: set colour values according ordering 
    scale_colour_manual(limits = c("Daily Stock Prices", "Rolling Mean", 
                                   "Loess Reg"),
                        values = c("blue", "red", "darkblue")) +
  theme(legend.position = "bottom")

ordnung <- 39
bmw <- filter(bmw, Date >= "2010-11-01" & Date <= "2011-11-30")
bmw %<>% mutate(rol_mean = 
         stats::filter(Tageskurs, filter = rep(1/ordnung, ordnung)))
bmw_2011 <- ggplot(bmw, aes(x = Date, y = Tageskurs)) +
  # geom_point(col = "blue", na.rm = TRUE) +
  geom_line(aes(col = "Daily Stock Prices"), linetype = "solid", na.rm = TRUE) +
  geom_line(aes(y = rol_mean, col = "Rolling Mean"), linetype = "solid", size = 1.3,
            na.rm = TRUE) +
  geom_smooth(aes(col = "Loess Reg"), method = "loess", size = 1, na.rm = TRUE) +
  labs(x = "Date", y = "Stock Price / Euro", col = "") +
  ggtitle("Dayly BMW Stock Prices (see Fahrmeir Abb. 14.9, p.519)",
          subtitle = paste("Period of Time: ", min(bmw$Date), "-", max(bmw$Date),
                           "; Rolling Mean: ", ordnung, "days")) +
  # limits: ordering of colours
	# values: set colour values according ordering 
    scale_colour_manual(limits = c("Daily Stock Prices", "Rolling Mean", 
                                   "Loess Reg"),
                        values = c("blue", "red", "darkblue")) +
  theme(legend.position = "bottom")

grid.arrange(bmw_2000_2015, bmw_2011)
```


### Pflanzen Garten Abb 14.3 p. 504 ###

```{r garten, echo = FALSE}
#### Pflanzen Garten Abb 14.3 p. 504 ###
(garten <- as_tibble(read.table("Data_Fahrmeir/Pflanzen_Garten.txt", header = TRUE)))

# garten <- filter(garten, Year < 1992)

garten_long <- garten %>%
  pivot_longer(cols = Jan:Dec,
               names_to = c("Month"),
               values_to = "Price")
# garten_long$Month <- factor(garten_long$Month, levels = month.abb)
# garten_long$month <- as.integer(garten_long$Month)
# for plots with time scale years and months 
# column $year.month with decimal year.month values needed

garten_long %<>%  
  unite("Year_Month", Year, Month, sep = "-", remove = TRUE) %>% 
  mutate(Year_Month = yearmonth(Year_Month))

filter(garten_long, as.integer(Year_Month) <= as.integer(yearmonth("1992 Feb")))

ggplot(garten_long, aes(x = Year_Month, y = Price)) +
  geom_line(col = "blue") +
  labs(x = "Year", y = "Price / Euro") +
  ggtitle("Monatlicher Preisindex für 'Pflanzen, Güter für die Gartenpflege'",
          subtitle = "Fahrmeir Abb. 14.3, p. 504")
```


```{r garten zerlegung, echo = FALSE, fig.width= 7, fig.height= 5}
start <- c(1991,1)
end <- c(1995,12)

garten.ts <- ts(garten_long$Price,
                    start=start,end=end,frequency=12)
plot(stl(window(garten.ts, start=start,end=end),
         s.window="per",t.window=600), # saison = periodic, trend = liner
     main = "Garten Preisindex Zeitreihenzerlegung - linearer Trend & starre Saisonkomponente")

plot(stl(window(garten.ts, start=start,end=end),
         s.window=27,t.window=29), 
     main = "Garten Preisindex Zeitreihenzerlegung - Trend & Saisonkomponente nicht starr")

plot(stlplus(window(garten.ts, start=start,end=end),
         s.window="per",t.window=600), 
     main = "Garten Preisindex Zeitreihenzerlegung - linearer Trend & starre Saisonkomponente")

plot(stlplus(window(garten.ts, start=start,end=end),
         s.window=27,t.window=29), 
     main = "Garten Preisindex Zeitreihenzerlegung - Trend & Saisonkomponente nicht starr")

```


### IFO Abb 14.4 p. 505

```{r ifo,  echo = FALSE}
#### IFO Abb 14.4 p. 505 ###
ifo <- as_tibble(read.table("Data_Fahrmeir/ifo_zeitreihen.txt", header = TRUE))

ifo %<>%  
  separate(Zeit, c("Month", "Year"),  
           sep = "[[:punct:]{/}]", extra = "warn") %>% 
  unite("Year_Month", Year, Month, sep = "-", remove = TRUE) %>% 
  mutate(Year_Month = yearmonth(Year_Month)) %>% 
  rename("Bauhauptgewerbe Auslastung" = Bauhauptgewerbe_Kapazitätsauslastung,
         "Gewerbliche Wirtschaft" = Gewerbliche_Wirtschaft_Gki) 

ifo_long <- ifo %>%
   pivot_longer(cols = c("Bauhauptgewerbe Auslastung", "Gewerbliche Wirtschaft"),
               names_to = "Geschäftsklima",
               values_to = "Gki")

ifo_long <- as_tsibble(ifo_long, key = Geschäftsklima, index = Year_Month) 
ifo_long

ggplot(ifo_long, aes(x = Year_Month, y = Gki, col = Geschäftsklima)) +
  geom_line() +
  labs(x = "Year", y = "Geschäftsklimaindex") +
  theme(legend.position = "bottom") +
  ggtitle("IFO Geschäftsklimaindex 1991 - 2014",
          subtitle = "Fahrmeir Abb. 14.4, p. 505")
```


```{r ifo zerlegung, echo = FALSE, fig.width= 7, fig.height= 5}
ifo.ts <- ts(ifo[ , "Gewerbliche Wirtschaft"],
                    start = c(1991,1), end = c(2014,12), frequency = 12)
plot(stl(window(ifo.ts[,1], start=c(1991,1),end=c(2014,12)),
         s.window="per",t.window=600), # saison = periodic, trend = liner
     main = "IFO Geschäftsklimaindex Zeitreihenzerlegung - linearer Trend & starre Saisonkomponente")

plot(stl(window(ifo.ts[,1], start=c(1991,1),end=c(2014,12)),
         s.window=27,t.window=29), 
     main = "IFO Geschäftsklimaindex Zeitreihenzerlegung - Trend & Saisonkomponente nicht starr")
```


# Chapter 14.6 Zeitreihenanalyse mit R (p. 524)

## Stickoxide - Zeitreihenzerlegung linearer Trend & starre Saisonkomponente

```{r Stickoxide, eval = TRUE, echo = FALSE, fig.width= 7, fig.height= 5}
###################### 14.6 Zeitreihenanalyse mit R p. 524 ###

daten <- read.delim("Data_Fahrmeir/luftschadstoffe.txt")
Stickoxide.ts <- ts(daten$Stickoxide,
                    start=c(1991,1),end=c(2016,2),frequency=12)
plot(stl(window(Stickoxide.ts, start=c(1996,1),end=c(2015,12)),
         s.window="per",t.window=600), # saison = periodic, trend = liner
     main = "Stickoxid-Zeitreihenzerlegung linearer Trend & starre Saisonkomponente")


plot(stl(window(Stickoxide.ts, start=c(1996,1),end=c(2015,12)),
         s.window=7,t.window=59), 
     main = "Stickoxid-Zeitreihenzerlegung Trend & Saisonkomponente nicht starr")
```

## BMW Stock Values from Internet

for instrument = " " symbols (e.g.  "^gdaxi", bmw.de", "sie.de") see:
   https://finance.yahoo.com/quote/%5EGDAXI/components/

Download quote options: "Open", "High", "Low", "Close", "Adjusted" and "Volume"

- Close:    Close price adjusted for splits
- Adjusted: Adjusted close price adjusted for both dividends and splits  
     Korrektur durch Zurückrechnung;
     Adjustierung/Anpassung eines Aktienkurses bzgl.
         Splits/Kapitalerhöhung(en), Dividends/Auszahlung von Dividenden o.ä

```{r bmw2 index plot, eval = FALSE, echo = FALSE}
bmw2 <- get.hist.quote(instrument="bmw.de",
                       start="2000-01-01", end="2015-06-03",
                       quote="Adjusted")

plot(bmw2, type="l",
     xlab="Zeitraum: 3.1.2000-3.6.2015",
     ylab="Tageskurse der BMW-Aktie")
index <- 1:length(bmw2)

lo <- loess(bmw2$Adjusted~index)
lines(zoo::index(bmw2), c(NA,predict(lo),NA)) 
#       zoo:: required due to package tsibble

```
   

```{r bmw2, eval = TRUE, echo = FALSE, message=FALSE}
bmw2 <- get.hist.quote(instrument="bmw.de",
                       start="2000-01-01", end="2015-06-03",
                       quote="Adjusted")

stock_data <- as_tibble(rownames_to_column(as.data.frame(bmw2), var = "Date"))
stock_data$Date <- as.Date(stock_data$Date)

ordnung <- 201
stock_data %<>% mutate(rol_mean =
         stats::filter(Adjusted, filter = rep(1/ordnung, ordnung)))

ggplot(stock_data, aes(x = Date, y = Adjusted)) +
  # geom_point(col = "blue", na.rm = TRUE) +
  geom_line(aes(y = Adjusted, col = "Daily Stock Prices"), linetype = "solid",
            na.rm = TRUE) +
  geom_line(aes(y = rol_mean, col = "Rolling Mean"), linetype = "solid",
            size = 1.3, na.rm = TRUE) +
  geom_smooth(aes(col = "Loess Reg"), method = "loess", size = 1, na.rm = TRUE) +
  labs(x = "Date", y = "stock Price / Euro", col = "") +
  ggtitle("Dayly BMW stock Prices (see Fahrmeir Abb. 14.9, p.519)",
          subtitle = paste("Period of Time: ", min(stock_data$Date), "-",
                           max(stock_data$Date), "; Rolling Mean: ",
                           ordnung, "days")) +
  # limits: ordering of colours
	# values: set colour values according ordering
    scale_colour_manual(limits = c("Daily Stock Prices", "Rolling Mean",
                                   "Loess Reg"),
                        values = c("blue", "red", "darkblue")) +
  theme(legend.position = "bottom")

# w/o end: = yesterday
# S&P 500 (^GSPC)
# x <- get.hist.quote(instrument = "^gspc", start = "1998-01-01",
#                     quote = "Close")
# plot(x)
```

## Stock Prices for different Quotes

```{r different Quotes, eval = TRUE, echo = FALSE, fig.width= 7, fig.height= 8}
# ^GDAXI  DAX PERFORMANCE-INDEX
# 
# DTE.DE	Deutsche Telekom AG
# DBK.DE	Deutsche Bank Aktiengesellschaft
# BMW.DE	Bayerische Motoren Werke Aktiengesellschaft
# MUV2.DE	Munchener Ruckversicherungs-Gesellschaft Aktiengesellschaft
# SIE.DE	Siemens Aktiengesellschaft

company <- "DBK.DE"
quote_all <- c("Open", "High", "Low", "Close", "Adjusted", "Volume")
stock_data <- get.hist.quote(instrument = company, 
                    start = "2010-01-01",
                    quote = quote_all)
# plot(stock_data, main = "BMW - Closing Value and Volume")

stock_data <- as_tibble(rownames_to_column(as.data.frame(stock_data), var = "Date"))
stock_data$Date <- as.Date(stock_data$Date)

stock_data_long <- stock_data %>%
   pivot_longer(cols = Open:Adjusted,
               names_to = c("Quote"),
               values_to = "Price")

ggplot(stock_data_long, aes(x = Date, Price, col = Quote)) +
    facet_wrap( ~ Quote, ncol = 1, scales = "free" ) +
    # geom_point(col = "blue", shape = 20, na.rm = TRUE) +
    geom_line() +
    theme(legend.position = "bottom") +
    labs(y = "Stock Price / €") +
    ggtitle("Stock Prices for different Quotes", 
            subtitle = paste("Company: ", company)) 
```

## Adjusted Stock Prices - DAX compared with Companies

```{r DAX_Companies, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=8}

# ^GDAXI  DAX PERFORMANCE-INDEX
#
# BMW.DE	Bayerische Motoren Werke AG
# DBK.DE	Deutsche Bank AG
# DTE.DE	Deutsche Telekom AG
# MUV2.DE	Munchener Ruckversicherungs-Gesellschaft AG
# SIE.DE	Siemens AG


company_tb <- tibble(Company =
                       c("DAX", "BMW", "Deutsche Bank", 
                         "Munich Re", "Siemens", "Deutsche Telekom"),
                     Symbol = c("^gdaxi",  "BMW.DE", "DBK.DE", 
                                "MUV2.DE", "SIE.DE", "DTE.DE"))

start_of_period <- "2010-01-01"
end_of_period <-   Sys.Date() - 1
stock_data_all <- NULL

for (i in seq_along(company_tb$Company )) {
  stock_data <- get.hist.quote(instrument = company_tb$Symbol[i],
                               start = start_of_period, end = end_of_period,
                               quote = c("Adj"))
  stock_data <- as_tibble(rownames_to_column(as.data.frame(stock_data), var = "Date"))
  stock_data$Date <- as.Date(stock_data$Date)
  stock_data$Company <- company_tb$Company[i]
  stock_data_all <- bind_rows(stock_data, stock_data_all)
}

stock_data_all$Company <-  factor(stock_data_all$Company,
                                  levels = company_tb$Company)

stock_plot <- ggplot(stock_data_all, aes(x = Date, Adjusted, col = Company)) +
  facet_wrap( ~ Company, ncol = 1, scales = "free" ) +
  # geom_point(shape = 20, na.rm = TRUE) +
  geom_line(na.rm = TRUE) +  # TRUE rows w/ NA are silently removed, no warning
  labs(y = "Adjusted Stock Price / €") +
  theme(legend.position = "none") +
  ggtitle("Adjusted Stock Prices - DAX compared with Companies", 
          subtitle = paste(start_of_period, "-", end_of_period))

stock_plot

# library(DT)
# DT::datatable(filter(stock_data_all, Date >= as.Date("2015-12-15") & Date <= as.Date("2015-12-31")))

stock_plot + geom_point(shape = 20, na.rm = TRUE) +
  xlim(as.Date("2015-12-01"), as.Date("2015-12-31")) 
# xlim => ylims von ganzem Datensatz
# Alternative: Datensatz filtern und uaf plot geben: stock_plot %+% filter(sto..)



```

## With tsibble - non trading days are filled with default NA

```{r DAX_w NAs, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=8}
stock_data_filter <- filter(stock_data_all, Date >= as.Date("2015-12-01")
                          & Date <= as.Date("2015-12-31"))
stock_data_tsbl <-
  as_tsibble(stock_data_filter, index = Date, key = Company)
stock_data_tsbl %>%  has_gaps()   # Check for gaps
stock_data_tsbl %>%  scan_gaps()  # Reveal
stock_data_tsbl %>%  count_gaps() # Summarise - number of gaps from - to
stock_data_tsbl %<>%  fill_gaps() # Fill in time gaps
stock_data_tsbl %>%  has_gaps()
key(stock_data_tsbl)
# slide()
# future_pstretch() {tsibble} Stretching window in parallel

# stock_data_tsbl %>% index_by(..) # group_by time index and collapse with summarise()
  
# tidy() model parameters
# glance() a one-row per model summary
# augment()
# forecast() {fabletools}

stock_data_tsbl <- 
  as_tsibble(stock_data_all, index = Date, key = Company) %>% 
  fill_gaps()  # fill gaps with NAs
# datatable(stock_data_tsbl)

stock_plot <- ggplot(stock_data_tsbl, aes(x = Date, Adjusted, col = Company)) +
  facet_wrap( ~ Company, ncol = 1, scales = "free" ) +
  geom_line(na.rm = TRUE) +  # TRUE rows w/ NA are silently removed, no warning
  labs(y = "Adjusted Stock Price / €") +
  theme(legend.position = "none") +
  ggtitle("Adjusted Stock Prices - DAX compared with Companies", 
          subtitle = paste(start_of_period, "-", end_of_period))
stock_plot
stock_plot %+% filter(stock_data_tsbl, Date >= as.Date("2015-12-15") 
                          & Date <= as.Date("2015-12-31")) +
  geom_point(shape = 20, na.rm = TRUE)
```

