---
title: "Time Series Analysis of Weather Data"
subtitle: "see Fahrmeier Statistik Chapter 14 'Zeitreihen'"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
output:
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 12
    number_sections: yes
    toc: yes
  html_document: default
  word_document:
    toc: yes
  odt_document:
urlcolor: blue
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r initialization, echo = FALSE, message = FALSE, warning = FALSE}
######  variables clean up, library() and	source() statements
setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis")
rm(list=ls()) # deletes all existing objects / variables !!

library(zoo)
library(tseries) # Finance data can be easily queried using the get.hist.quote
library(stlplus) # Enhanced Seasonal Decomposition of Time Series by Loess
library(tidyverse)
library(magrittr)
library(tsibble)
library(lubridate)  # for year(date) function
# for package overview see:
#                   https://cran.r-project.org/web/views/TimeSeries.html
#                   https://slides.earo.me/rstudioconf19/#16
library(DT)     # R interface to the JavaScript library DataTables
library(fable)
# vignette("fable")
library(gridExtra)

# packages to be considered:
# library(forecast)
# library(xts)

#################################
# library(devtools)
# # devtools::load_all("animals4") # loads package incl. data files (if available)
# 
# devtools::install("D:/Wolfgang/Programs-R/statistikv8")
# devtools::load_all("D:/Wolfgang/Programs-R/statistikv8")
#                             # loads package incl. data files (if available)
# library("statistikv8")
# help("abb14_loess")
# dir("D:/Wolfgang/Programs-R/statistikv8/R/", full.names = TRUE)
#################################

source("./uts_TimeSeries.R")  # utility functions for time series
source("./ggts_TimeSeries.R")

```

```{r select city}
city_list <- c("Cottbus", "Giessen", "Mannheim", "Basel", "England")
city <- city_list[3]   # city/country of data
```


```{r ggplot settings, echo = FALSE, message = FALSE, warning = FALSE}

# theme_set(theme_grey())
theme_replace(plot.title = element_text(hjust = 0.5, face = "bold.italic", 
                                        color = "darkcyan", size = 12),
              plot.subtitle = element_text(hjust = 0.5, face = "italic", 
                                           color = "darkcyan"))
```

# Time Series Analysis

## Overview of Time Series Data - Plot and add Rolling Mean 

```{r get weather data, echo = FALSE, message = FALSE}

# weather_data <- readRDS(paste0("../R-WeatherAnalysis/WeatherData_", city, ".rds"))

weather_data <- read_csv(paste0(
    "../R-WeatherAnalysis/Data_Weather/Weather-Data-Input_", city, ".csv"))
# saveRDS(weather_data, paste0("../R-WeatherAnalysis/WeatherData_", city, ".rds"))


# Read File Format:
# weather_data
# Year Measure   Jan  ... Dec Year_avg Winter_avg Spring_avg Summer_avg Fall_avg
# <dbl> <chr>       <dbl> ..<dbl>    <dbl>      <dbl>      <dbl>      <dbl>    <dbl>
# 1864 Temperature  -5.5  ...-2.2     7.7      NA           9.1        16.3     8.07
# :
# 2019 Precipitation  22.2...50.9     65.5      49.7        60.9       93.0     74.4
```

```{r data view, eval=FALSE, echo=FALSE}
## load data
# data("Melanoma", package="MASS")

# interactive data overview; not working with knit => eval=FALSE
# DT::datatable(Melanoma, options=list(pageLength=6)) #interactive data overview
DT::datatable(weather_data, options=list(pageLength=14)) #interactive data overview
```


### Weather Data - Plot with Rolling Mean

```{r weather, echo = FALSE, fig.width= 7, fig.height= 4}

weather <- filter(weather_data, Measure == "Temperature")
# weather <- filter(weather, Year < 1992)

# weather <- filter(weather, Year >= 1890)


## generate test data with missing row (Year 149 deleted) and 
##                                     lot, 3, 2, 1 NA's in a line
test <- FALSE
if (test) {
  weather_test <- filter(weather, 
                         Year >= 1940 & Year <= 1950 & Year & Year != 1949)
  weather_test[1, 9:11] <- NA
  weather_test[2, 6:7] <- NA
  weather_test[2, 12] <- NA
  weather_test
  weather <- weather_test 
}

                       
weather_long <- weather %>%
  pivot_longer(cols = Jan:Dec,
               names_to = c("Month"),
               values_to = "Temperature")
# weather_long$Month <- factor(weather_long$Month, levels = month.abb)
# weather_long$month <- as.integer(weather_long$Month)
# for plots with time scale years and months 
# column $year.month with decimal year.month values needed

weather_long %<>%  
  unite("Year_Month", Year, Month, sep = "-", remove = TRUE) %>% 
  mutate(Year_Month = yearmonth(Year_Month))

span <- 15*12 # width of the rolling window
weather_plot <- 
  ggts_w_rol_mean(weather_long, Year_Month, Temperature, span = span)

y_temp <- expression(paste("Temperature /", ~degree*C)) 
weather_plot +  labs(y = y_temp) + ggtitle("Monthly Average Temperatures")

first_full_year <- min(filter(weather, !is.na(Jan))$Year)
last_full_year <- max(filter(weather, !is.na(Dec))$Year)

weather_plot +  labs(y = y_temp) + ggtitle("Monthly Average Temperatures") +
  xlim(yearmonth(last_full_year-30), yearmonth(last_full_year))
  # xlim(as.Date("1940-01-01"), as.Date("1950-12-31"))
  
```

## Time Series Decomposition and Seasonal Adjustment

```{r weather prep zerlegung, echo = FALSE, fig.width= 7, fig.height= 10}


st_year <- first_full_year
st_month <- 1
end_year <- last_full_year
end_month <- 12
freq <- 12

# ts objects don't allow missing years
weather_long <- as_tsibble(weather_long) %>% fill_gaps()

data_length <- (end_year - st_year) * freq + 
  (freq - st_month + 1) + (freq - end_month)

start_data <- as.Date(paste(st_year, "-", st_month, "-01", sep = ""))
end_data <- as.Date(paste(end_year, "-", end_month, "-01", sep = ""))

start_ts <- c(st_year, st_month)
end_ts <- c(end_year, end_month)

############ ts() !! attention ###########
# !! date Spalte imd df wird prinzipiell nicht angeschaut,
#    es wird einfach durch-indiziert und entsprechend den start & end Werten
#    die rownames und colnames und frequency benamst
#   => Datenanfang musss immer mit start Werten übereinstimmen !! 
#      muss aber nicht Anang / Ende des Jahres und ganze Saison sein
#   => es dürfen keine Jahre einfach ausgelassen sein => Jahre mit NA's einfügen   
#   => Frequncy/Saison muss zu den Zeitscheiben (monatlich/täglich/Quartale) 
#      passen ! 
data_filter <- 
  filter(weather_long, Year_Month >= start_data & Year_Month <= end_data)
if(nrow(data_filter) != data_length)
  stop("Data length = ", data_length, 
       " due to st_year = ", st_year, " st_month = ", st_month, 
       " end_year = ", end_year, " end_month = ", end_month,
       "\n differs to nrow(Data file) = ", nrow(data_filter),
       "\n => check missing years in data file and start/end settings")
data_ts <- ts(data_filter$Temperature, 
                 start=start_ts, end=end_ts, frequency = freq)
if(nrow(data_filter) != length(data_ts))
  stop("nrow(Data file) = ", nrow(data_filter),
       " differs to length(ts object) = ", length(data_ts),
       "\n => check missing years in data file and start/end settings",
       "\n => in 'ts object' data will be periodical replicated !")

```


```{r weather zerlegung, echo = FALSE, fig.width= 7, fig.height= 10}

############ stlplus()
full_range <- paste0(first_full_year, "-", last_full_year)
plot_title <- paste("Weather", 
                    "Time Series Decompression w/ Means of", full_range)

# stlplus_data <- stlplus(data_ts, s.window=1001, t.window=1201)
stlplus_data <- stlplus(data_ts, s.window="per", t.window=1201)
# plot(stlplus_data,      
#      main = "weather Zeitreihenzerlegung - linearer Trend & starre Saisonkomponente")
# gg_plot_and_data <- ggplot_stlplus_w_data(stlplus_data)


data_monthly_stlplus <- uts_stlplus_as_tibble(stlplus_data)

decomp_plot_l <- ggts_decomp(data_monthly_stlplus) +
    labs(x = "Year", y = y_temp, col = "") +
    labs(title = plot_title,
         subtitle = "w/ Periodic season & ~linear Trend in span of loess window")
decomp_plot_l

stlplus_data <- stlplus(data_ts, s.window=27, t.window=1201)

data_monthly_stlplus <- uts_stlplus_as_tibble(stlplus_data)

decomp_plot <- ggts_decomp(data_monthly_stlplus) +
    labs(x = "Year", y = y_temp, col = "") +
    labs(title = plot_title,
            subtitle = "w/ no regid Season and Trend Component")
decomp_plot

##  plot time window 

decomp_plot_l + xlim(yearmonth(last_full_year-30), yearmonth(last_full_year))

decomp_plot + xlim(yearmonth(last_full_year-30), yearmonth(last_full_year))

```


# ZeitreihenAbdalla

see YouTube Tutorials "youtube time series analysis with R"
## Plot Time Series

```{r Plot Zeitreiehe }

start_data <- as.Date("2001-01-01")
end_data <- as.Date("2019-12-31")
data_filter <- 
  filter(weather_long, Year_Month >= start_data & Year_Month <= end_data)

start_ts <- c(year(start_data), month(start_data))
end_ts <- c(year(end_data), month(end_data))

data_ts <- ts(data_filter$Temperature, 
                 start=start_ts, end=end_ts, frequency = freq)
plot(data_ts)

weather_plot <- 
  ggts_w_rol_mean(data_filter, Year_Month, Temperature, span = 36)
weather_plot

```

## Stationary Process Test

Strict-sense stationarity / Weak (wide-sense) stationarity

Trend Stationary - underlying trend (function solely of time) can be removed, 
leaving a stationary process

```{r}
## Stationarity Test
library(aTSA)     # for stationary.test(data_ts) and adf.test(data_ts) - same
library(fractal)  # for stationaryty()
library(LaplacesDemon)

stationary.test(data_ts)
adf.test(data_ts)

statio <-
  stationarity(data_filter$Temperature, n.taper = 6,
               n.block = max(c(12, floor(logb(length(data_ts), base = 12)))),
               significance = 0.05, center = TRUE, recenter = FALSE)
summary(statio)

is.stationary(data_filter$Temperature)
# statcheck() from Schlittgen
source("./Data_Schlittgen/tsutil.R")
statcheck(data_filter$Temperature, 5)

data_temp <- dplyr::select(weather, Year, Jan:Dec)
lm(data_temp)
summary(lm(data_temp))

model_lm <- lm(Jan ~ Feb + Mar^2 + Apr^3 + Mar^4, data_temp)
summary(model_lm)

##########################

data_temp <- dplyr::select(weather, Year, Year_avg)
model_lm <- lm(Year_avg ~ Year, data_temp)
summary(model_lm)

model_pol_lm <- lm(Year_avg ~ Year + Year^2 + Year^3, data_temp)
summary(model_pol_lm)
```

## Periodicities - Season Frequency  

### Correlogram - autocovariance / autocorrelation function

```{r}
# Korrelogramm
# „empirische Autokovarianz- und Autokorrelationsfunktion (ACF)
# 
#  Lag - Verschiebung, des betrachteten Zeitfensters
#  Lag 0: Jan mit Jan (sich selbst)
#  Lag 1: Jan mit Feb (next Feb , einen Monat weiter)
#  
#  Lag 12: Jan mit Jan (des nächsten Jahres)
#  t <- c(1,2,3) => lag(t): [1] NA  1  2
data_acf <- acf(data_ts, lag.max = 24) # type = "correlation" (the default)
# plot(data_acf)

# Temperaturzeitreihe weist natürliche saisonale Schwankungen auf
# => gut im Korellogramm zu erkennen kann
# Der Januar wird als Basis gewählt und die Monate um den Januar herum weisen 
#   eine sehr hohe positive Korrelation auf
# die Sommermonate in der Mitte (0,5) des Jahres weisen eine starke
#                  negative Korrelation auf
data_acf_tbl <- enframe(as.vector(data_acf$acf), name = NULL) %>% 
  mutate(ACF = as.vector(value),
         Lag = as.vector(data_acf$lag))
ggplot(data_acf_tbl, aes(Lag, ACF)) +
  geom_bar(stat = "identity", width = 0.01) +
    ggtitle( paste("Time Series Correlogram - Type ", data_acf$type))

```

### Periodogram - Estimate Spectral Density of a Time Series

```{r}
# Periodogramm
# Das Periodogramm ist das Pendant zum Korellogramm auf Spektralebene
# Es ist die abgebildete Fouriertransformation der Autokovarianzfunktion 
data_period <- spec.pgram(data_ts, taper = 0, pad = 0, fast = TRUE, demean = FALSE, 
           detrend = TRUE, plot = TRUE, na.action = na.fail)
#  Frequenz λ = 1: signiﬁkanter Ausschlag stattﬁndet
#  => vorliegende Muster wiederholt sich jede volle Frequenz / jedes Jahr
#  Die restlichen Ausschläge sind bereits zu zufällig, um sie signiﬁkant 
#                                                    einordnen zu können.

data_period_tbl <- as_tibble(data_period$freq) %>% 
  mutate(Frequency = value,
         Spectrum = data_period$spec)
ggplot(data_period_tbl, aes(Frequency, Spectrum)) +
  geom_line() +
  labs(x = paste("Frequency\n", "bandwidth = ", data_period$bandwidth)) +
  scale_y_log10() +
    ggtitle( paste("Time Series ", data_period$method))
```

## Seasonal Decomposition - Seasonal and Trend extraction of Time Series

Check remainder for white noise

```{r}
data_decomp <- decompose(data_ts)
plot(data_decomp)

############ stlplus()
stlplus_data <- stlplus(data_ts, s.window= 13, t.window=21)
# plot(stlplus_data,      
#      main = "weather Zeitreihenzerlegung - linearer Trend & starre Saisonkomponente")

data_monthly_stlplus <- uts_stlplus_as_tibble(stlplus_data)

decomp_plot <- ggts_decomp(data_monthly_stlplus) +
    labs(x = "Year", y = y_temp, col = "") +
    ggtitle("Weather Time Series Decompression",
            subtitle = "w/ Season & Trend in span of loess window")
decomp_plot
```

### Saisonbereinigung mithilfe von Filtern

```{r }
freq <- 12
data_temperature_values <- as.vector(data_ts)

data_rol_mean_year <- stats::filter(data_temperature_values, 
                                    rep(1/freq, freq), method = "convolution") 
# default method
data_remainder <- data_temperature_values - data_rol_mean_year 
par(mfrow = c(3, 1))
plot(data_temperature_values, type = "l")
plot(data_rol_mean_year, type = "l")
plot(data_remainder, type = "l")

## Vergleich Alternative dieses Befehls ist die recursive Methode.
data_rol_mean_year <- stats::filter(data_temperature_values, rep(1/freq, freq), 
                                    method = "recursive")
data_remainder <- data_temperature_values - data_rol_mean_year 
par(mfrow = c(3, 1))
plot(data_temperature_values, type = "l")
plot(data_rol_mean_year, type = "l", main = "recursive")
plot(data_remainder, type = "l")


data_filter %<>% mutate(rol_mean_year = 
                          as.numeric(stats::filter(Temperature,
                                                   rep(1/freq, freq), method = "convolution")),
                        remainder = Temperature - rol_mean_year)

data_filter_long <- data_filter %>%
  pivot_longer(cols = c("Temperature", "rol_mean_year", "remainder"),
               names_to = c("Data"),
               values_to = "count") %>% 
  mutate(Data = factor(Data, levels = c("Temperature", 
                                        "rol_mean_year", "remainder")))

gg_plot <-  ggplot(data_filter_long, aes(Year_Month, count, col = Data)) +
  facet_wrap( ~ Data, ncol = 1, scales = "free" ) +
  geom_line(na.rm = TRUE) +  # # TRUE rows w/ NA are silently removed, no warning
  theme(legend.position = "bottom")  +
  labs(x = "Year", col = "") +
  ggtitle("Time Series Decompression",
          subtitle = "with Function stlplus()")
gg_plot
```

### Partial autocorrelation function (PACF)


# Backup for Tests 

```{r ggplot test, eval = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
### ggplot function definitions
# plot_survival_split()
# 
 
# plot with running/rolling mean - attention! use aes_(aes with underscore)
plot_test_roll_mean <- function(data, x, y, span) {
  
  # new <- data %>% mutate(neu = as.name(!!(y)))
  # print(new)
  gg_plot <- ggplot(data, aes_(x = substitute(x), y = substitute(y))) +
    geom_line(aes(col = "Measured Values")) +
    geom_line(aes(y = rol_mean, col = "Rolling Mean"), na.rm = TRUE) +
    geom_smooth(aes(col = "Loess Reg"), method = "loess", size = 1, na.rm = TRUE) +
    # values: set colour values according ordering
    scale_colour_manual(limits = 
                          c("Measured Values", "Rolling Mean", "Loess Reg"),
                        values = c("blue", "red", "darkblue")) +
    theme(legend.position = "bottom") +
    labs(x = "Year", y = substitute(y), col = "") +
    ggtitle("Monthly Average Temperatures",
            subtitle = paste("with Rolling Mean over", span, 
                             "measured values and Loess Regression"))
  return(gg_plot)
}

# plot_test_roll_mean (weather_long, Year_Month, Temperature, span = span)
# mauna_co2_plot <-
#   plot_test_roll_mean (mauna_loa_co2_long, 
#                            Year_Month, interpolated, span = span)
# 
# y_co2 <- expression(paste(CO[2], " Concentration / ppm")) 
# mauna_co2_plot +
#   labs(y = y_co2) +
#   ggtitle("Mauna Loa - Monthly Average CO2 Concentration")


```

