
```{r}
city_list <- c("Cottbus", "Giessen", "Hohenpeissenberg", "Potsdam",
               "Basel", "Davos", "England", "Mauna Loa", "Basel_Giessen")
# city <- city_list[1]   # city/country of data setting
city <- c("Mauna Loa")

slice(data_monthly, 1:6)
slice(data_yearly, 1:6)

# saveRDS(data_monthly, paste0("test_data_monthly_", city, ".rds"))
# saveRDS(data_yearly, paste0("test_data_yearly_", city, ".rds"))

data_monthly <- readRDS(paste0("test_data_monthly_", city, ".rds"))
data_yearly <- readRDS(paste0("test_data_yearly_", city, ".rds"))


```


```{r}



```

```{r}

## !! for data with more than one time series
## => filter to get single time series needed for
##    - decomposition with stlplus() 
##    - usage of  e.g. gg_lag(), gg_tsdisplay()
##
## => will be ensured that in "dataset" always to keys exist (City, Measure)
##

# so far:   key <- c("Measure", "City") => to be changed c("City", "Measure")
# if ordering needed
# City_A Measure_A
#             City_A Measure_B
#             City_B Measure_A
#
# enssure right ordering rsp. change if needed
data_monthly %<>% as_tsibble(index = Year_Month, key = c("City", "Measure"))


key_data(data_monthly) # tibble with key name(s) and subnames
key(data_monthly)      # provides "group" key name;
key_vars(data_monthly) # provides the names of the keys
n_keys(data_monthly)   # number of separate time series

if (n_keys(data_monthly) == 1 | length(key(data_monthly)) == 2) {  
  for (i in unique(key_data(data_monthly)[[1]])) {
    for (j in unique(key_data(data_monthly)[[2]])) {
      print(paste("First Key:", names(key_data(data_monthly)[1])))
      print(paste("Second Key:", names(key_data(data_monthly)[2])))
      print(paste(i, "Time Series of", j))
    }
  }
} else {
  abort(paste("nothing implemented for n_keys/key_data/key(data_monthly) =", 
              n_keys(data_monthly), key_data(data_monthly), (key(data_monthly))))
}

```



# Analysis




## Forecasting with ETS models

```{r Forecasting with ETS models, eval = TRUE, fig.width= 7, fig.height= 4}

fit <- filter(data_monthly, year(Year_Month) <= 2000) %>%  
  model(ETS(count))
fit <- filter(data_monthly, year(Year_Month) <= 2000) %>%  
  model(ETS(count ~ season("A")))

glance(fit)
report(fit)
fit$"ETS(count)"
model_ets <- fit[[3]] #  fit$"ETS(count)"
model_ets <- as.character(format(model_ets))

fit <- data_monthly %>%
  model(ETS_ANA = ETS(count ~ error("A") + trend("N") + season("A")),
        ETS_AAA = ETS(count ~ error("A") + trend("A") + season("A")),
        ETS_AAdA = ETS(count ~ error("A") + trend("Ad") + season("A")),
        ETS_AMA = ETS(count ~ error("A") + trend("M") + season("A")),
        ETS_MNA = ETS(count ~ error("M") + trend("N") + season("A")),
        ETS_MAA = ETS(count ~ error("M") + trend("A") + season("A")),
        ETS_MAdA = ETS(count ~ error("M") + trend("Ad") + season("A")),
        ETS_MMA = ETS(count ~ error("M") + trend("M") + season("A")))
# model_ets <- fit$ETS_AAA
# model_ets <- as.character(format(model_ets))

glance(fit)
glance(fit) %>% summarise(min(AIC), min(AICc), min(BIC))
# report(fit) # only working for single modles !
print("For best model choose model with lowest AIC, AICc, BIC")

fc <- fit %>% fabletools::forecast(h = "10 years")

# fc_filter <- filter(fc, .model == "ETS_AAA" & .model == "ETS_MAA")
# w/ filter scale easily min of xlim
data <- filter(data_monthly, year(Year_Month) >= 2010) 
fc_filter %>%
  autoplot(data, level = 95, alpha = 0.5) +
  # autoplot: data + forecast AAA & forecast MAA w/ or w/o signif. level 
  #           (prediction interval);  level = NULL rsp level = 95 (=95%))
  labs( x = "Year", y = y_label,
        title = "Forecasts from ETS models", subtitle = model_ets) +
  guides(colour = guide_legend(title = "Forecast"))



# print(getAnywhere(report.ETS))  # method.class => provides source code
report(filter(fit, Temp_Precip == "Temperature" & City == "Basel"))
report(filter(fit, Temp_Precip == "Precipitation" & City == "Basel"))

tidy(fit)    # summarizes a model's statistical findings such as coefficients
augment(fit) # adds columns to the original data such as predictions, residuals 
             #                     and cluster assignments
glance(fit)  # provides a one-row summary of model-level statistics.

components(fit)

# plot of ETS components
components(fit) %>%
  autoplot() +
  ggtitle(paste(model_ets, "Model - Components"))

residuals(fit)
residuals(fit, type = "response")

# plot forecast next "4 years"
fit %>% fabletools::forecast(h = "4 years") %>%
  autoplot() +
  ylab(y_label) +
  ggtitle("Forecast next 4 years", subtitle = topic)

# plot forecast next "4 years" with previous data
fit %>% fabletools::forecast(h = "4 years") %>%
  autoplot(data_monthly) +  # adding previous data = data_monthly, before forecast
  ylab(y_label) +
  ggtitle("Measured Data with Forecast next 4 years", subtitle = topic)

fit %>% fabletools::forecast(h = "30 years") %>%
  autoplot(data_monthly) +  # adding measured data before forecast
  ylab(y_label) +
  ggtitle("Measured Data with Forecast next 30 years", subtitle = topic)


# -  a correlogram of the forecast errors 
# acf(rainseriesforecasts2$residuals, lag.max=20)

acf(residuals(fit), lag.max=48)
pacf(residuals(fit), lag.max=48)

# - test significant evidence for non-zero correlations at lags 1-20
# => carry out a Ljung-Box test
# Box.test(rainseriesforecasts2$residuals, lag=20, type="Ljung-Box")
  #   Box-Ljung test
  # data:  rainseriesforecasts2$residuals
  # X-squared = 17.4008, df = 20, p-value = 0.6268
Box.test(residuals(fit)$.resid, lag=48,, type="Ljung-Box")
cat("Null Hypothesis of independence in a given time series - for p < 0.05: reject H_0")
# - plot residuals
# plot.ts(rainseriesforecasts2$residuals)
# 
# - check whether the forecast errors are normally distributed with mean zero
# plot a histogram of the forecast errors, with an overlaid normal curve with 
# same mean zero and standard deviation
ggts_histo_forecast_resid(residuals(fit)) # make a histogram
# shows the plot that the distribution of forecast errors is roughly
# centred on zero, and is more or less normally ?

# same results with Holt_Winter Verahren und forecast::ets() Verfahren
# siehe data Test_Holt_Winter_verfahren
# library(forecast)
# data_ets <- ets(data_ts)
# plot(data_ets)
# plot(forecast::forecast(data_ets))
# plot(forecast::forecast(data_ets, h = 30*12))  # predict 360 data points = 30 years
# predict(data_ets, n.ahead = 2, prediction.interval = TRUE)
# 
# acf(data_ets$residuals, lag.max=48)
# pacf(data_ets$residuals, lag.max=48)
# 
# Box.test(data_ets$residuals, lag=48,, type="Ljung-Box")
# plot.ts(data_ets$residuals)
# plotForecastErrors(data_ets$residuals)
# 
# 
```


# Backup Alternative Plots



```{r covered tsibble plots, eval = FALSE}

data <- data_monthly   # %>% dplyr::select(Year_Month, count) 

## taken over #####################
data_monthly %>% gg_season(count, labels = "none") +
  labs(x = "Year",   y = y_label, 
       title =expression(paste("Mauna Loa - Yearly ", CO[2], 
                               "Time Plots over Months")))

data_monthly %>% gg_season(count, labels = "none", period = "6*year") +
  labs( x = "Year", y = "CO2 / ppm") +
  ggtitle("Seasonal plot: CO2 Concentration from Mauna Loa")



data %>% gg_season(count, period = "15*year", labels = "right") +
  labs( x = "Year", y = "CO2 / ppm") +
  ggtitle("Seasonal plot: CO2 Concentration from Mauna Loa")

data %>% gg_subseries(count) +
  labs( x = "Year", y = "CO2 / ppm") +
  ggtitle("Seasonal subseries plot: CO2 Concentration from Mauna Loa")

data %>% gg_tsdisplay(count) +  
  labs( x = "Year", y = "CO2 / ppm") +
ggtitle("CO2 Concentration - Time Series, ACF and Season Plot")



```



## Stationary Process Test

Strict-sense stationarity / Weak (wide-sense) stationarity

Trend Stationary - underlying trend (function solely of time) can be removed, 
leaving a stationary process

```{r eval = TRUE}
## Stationarity Test
library(aTSA)     # for stationary.test(data_ts) and adf.test(data_ts) - same
library(fractal)  # for stationaryty()
library(LaplacesDemon)

freq <- 12
data <- data_monthly_stlplus  # w/ numeric "Time"
# as.numeric(as.Date("2018-09-24") - as.Date("2017-10-24"))
# update data_ts with replaced NAs, e.g. adf.test() does not allow NAs
data_ts <- ts(data$count, start=c(first_year, 1), frequency = freq)

stationary.test(data_ts)
adf.test(data_ts)

statio <-
  stationarity(data_monthly$count, n.taper = 6,
               n.block = max(c(12, floor(logb(length(data_ts), base = 12)))),
               significance = 0.05, center = TRUE, recenter = FALSE)
summary(statio)

is.stationary(data_monthly$count)
# statcheck() from Schlittgen
source("./Data_Schlittgen/tsutil.R")
aspectratio(data_ts)  # für plot, Breite zu Höhe 
statcheck(data$count, 5) #Berechnung der deskriptiven Maßzahlen ist elementar. 
     # Um die Kovarianzstationarität zu überprüfen, wird die Funktion eingesetzt

vartable(data_ts,12) # variate Differenzen weisen darauf hin, dass einmal 
                     # einfache und einmal saisonale Differenzen zu bilden sind
                     # sind, um Stationarität zu erreichen.

test <- FALSE
if (test) {
  data_monthly_wide <- 
    readRDS(paste0("../R-WeatherAnalysis/WeatherData_", "Cottbus", ".rds"))
  weather <- filter(data_monthly_wide, Temp_Precip == "Temperature")
  test_monthly_wide <- filter(weather, Year >= 1940 & Year <= 1954)
  # note:   & Year & Year != 1949 not feasible, otherwise ts() indexing wrong
  test_monthly_wide[1, 9:11] <- NA
  test_monthly_wide[2, 6:7] <- NA
  test_monthly_wide[2, 12] <- NA
  test_monthly_wide
  rm(data_monthly)
  rm(data_ts)
  data_monthly <- test_monthly_wide %>% 
    pivot_longer(cols = Jan:Dec,
                 names_to = c("Month"),
                 values_to = "count") %>% 
    dplyr::select(Year, Month, count)
  first_year <- min(data_monthly$Year)
  data_ts <- ts(data_monthly$count, start=c(first_year, 1), frequency = freq)
  
  missar(data_ts, p=3, 1) 
  # AR method: missing values in a time series by conditional exspectations of AR(p)
  data_ts <- missls(data_ts, p=3, 1) 
  # KQ method: Minimum Mean Square Error Interpolator to ﬁll missings using LS approach
}


# bandfilt(data_ts, 7, 12, 24)
# View(data_ts)
# 
# (model_lm_1 <- lm(count ~ Year_Month, data))
# summary(model_lm_1)
# plot(model_lm_1)
# 
# model_lm_2 <- lm(count ~ Year_Month + Year_Month^2, data)
# summary(model_lm_2)
# plot(model_lm_2)


```


```{r Beispiel 10.10, eval = FALSE}
library(dlm)                                                                                 
# co2 <- scan("./Data_Schlittgen/schauinsland.dat")  
co2 <- data_ts  # co2 <- ts(co2,start = 2005,frequency=12)
dlmco <- dlmModPoly(2) + dlmModSeas(12)
m1 <- c(382,0.1,rep(0,11))  
c1 <- diag(c( 0.1, 0.1, rep(100,11)))

buildFun <- function(x) { 
  W(dlmco)[1:3,1:3] <- diag(exp(x[1:3]))
  V(dlmco) <- exp(x[4]) 
  C0(dlmco) <- c1
  m0(dlmco) <- m1 
  return(dlmco)
}              

fit <- dlmMLE(co2, parm=rep(1,4), build=buildFun)
fit$conv
dlm.co2 <- buildFun(fit$par)                                              
coFilter <- dlmFilter(co2, mod=dlm.co2)
coSmooth <- dlmSmooth(coFilter) 

plot(co2,type="o") 
lines(dropFirst(coFilter$m[,1]))
lines(dropFirst(coSmooth$s[,1]))
plot(dropFirst(coSmooth$s[,3]))

fut1 <- dlmForecast(coFilter, n=12)
x <- cbind(co2,fut1$f,fut1$f-1.96*sqrt(unlist(fut1$Q)),
           fut1$f+1.96*sqrt(unlist(fut1$Q)))                                
plot.ts(x,plot.type="s",lty=c(1,1,2,2)) 
points(2005+28/12,co2[29])




```





# Backup

## `r title_rmd` - Data Source 

**National Oceanic & Atmospheric Administration - Earth System Research Laboratory**

*NOAA ESRL *<https://www.esrl.noaa.gov/gmd/ccgg/trends/global.html>

Data file: Mauna Loa CO2 monthly mean data

<https://www.esrl.noaa.gov/gmd/ccgg/trends/data.html>


## `r topic` - Average Yearly and Seasonal Data

```{r data print, eval = TRUE}

caption_CO2 <- y_label
knitr::kable(data_yearly, digits = 1,
             caption = paste("Yearly", caption_CO2))


# knitr::kable(yearly_data_period_wide, digits = 1, 
#              caption = paste(span, "-years Periods (last period year given) Data -  ", 
#                              caption_CO2 ))

```

```{r end, echo = FALSE}
###########################################################################
Prog.End <- Sys.time()
run.time <- round((Prog.End - Prog.Start), digits = 2)
message("Program executed on: ", Prog.End, "\t Program run time: ", run.time, " secs")
```
