---
title: Coronavirus Covid-19 / SARS-CoV-2 
subtitle: Time Series of Confirmed, Death and Recovered Cases
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 12
    number_sections: yes
    toc: yes
    toc_depth: 3
urlcolor: blue
papersize: a4

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 7, fig.height = 4)

```


```{r initialization, eval = TRUE, include = FALSE}
#           fig.width=12, fig.height=6 ## old: fig.width=14.4, fig.height=7.2
################################################################################
## Plot and Analyze                                                           ##
## Corona Virus -Time Series of Confirmed, Death and Recovered Cases          ##
##                                                                            ##
##                                              Wolfgang Vollmer, March 2020  ##
##                                                                            ##
################################################################################

######  variables clean up, library() and	source() statements
# setwd("D:/Wolfgang/Programs-R/R-WeatherAnalysis")
setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis/Corona-Virus")
rm(list=ls()) # deletes all existing objects / variables !!
Prog.Start <- Sys.time()

library(tidyverse)
library(magrittr)
library(lubridate)
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(stlplus) # Enhanced Seasonal Decomposition of Time Series by Loess
library(DT)      # R interface to the JavaScript library DataTables
library(rlang)
library(gridExtra)

library(janitor)
library(highcharter)


######  Plot and util functions
# source("./../uts_TimeSeries.R")  # utility functions for time series
# source("./../ggts_TimeSeries.R") # ggplot2 functions for time series plots

theme_replace(
  plot.title = element_text(
    hjust = 0.5, face = "bold.italic", color = "darkcyan", size = 12),
  plot.subtitle = element_text(
    hjust = 0.5, face = "italic", color = "darkcyan")
)
x_axis_theme <- element_text(size = 14)
```

# Source data file: covid_19_data.csv

Downloaded from <https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/>


```{r read data bind rows and chanfge to long format}

# source data file: covid_19_data.csv from
# <https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/>

corona_confirmed <- read_csv(
  "./novel-corona-virus-2019-dataset/time_series_covid_19_confirmed.csv") %>% 
  mutate(count_type = "Confirmed")
corona_deaths <- read_csv(
  "./novel-corona-virus-2019-dataset/time_series_covid_19_deaths.csv") %>% 
  mutate(count_type = "Deaths")
corona_recovered <- read_csv(
  "./novel-corona-virus-2019-dataset/time_series_covid_19_recovered.csv")  %>% 
  mutate(count_type = "Recovered")

sum((corona_confirmed %>% filter(`Country/Region` == "US"))$`3/10/20`, na.rm = TRUE)
#> [1] 1670
# sum((corona_confirmed %>% filter(`Country/Region` == "US"))$`3/11/20`, na.rm = TRUE)
#> [1] 1281

corona_wide <- bind_rows(corona_confirmed, corona_deaths, corona_recovered)

corona <- corona_wide  %>%
    pivot_longer(cols = c(-`Province/State`, -`Country/Region`, -count_type, 
                          -Lat,  -Long ), 
                 names_to = c("Date"),
                 values_to = "count")

corona %<>% mutate(Date = as.Date(Date, format = "%m/%d/%y"))
corona %<>% rename(Country = `Country/Region`) 

```


```{r data clean-up}

# clean-up column names
# corona %>% janitor::clean_names()


# clean-up content of column Country
corona %<>% mutate(
  Country = case_when(
  Country == "Mainland China" ~ "China",
  Country == "Hong Kong SAR" ~ "Hong Kong",
  Country == "Iran (Islamic Republic of)" ~ "Iran",
  Country == "Macao SAR" ~ "Macao",
  Country == "occupied Palestinian territory" ~ "Palestine",
  Country == "Republic of Korea" ~ "South Korea",
  Country == "Korea, South" ~ "South Korea",
  Country == "Taipei and environs" ~ "Taiwan",
  Country == "US" ~ "United States of America",
  Country == "Viet Nam" ~ "Vietnam",
  TRUE ~ as.character(Country)))

# clean-up content of column `Province/State`
# replace NAs by Country
# replace "Bavaria"  by  Country since `Province/State` exists only until 
# end of January, as begin of February inluded in "Germany" data
# `Province/State` == "Bavaria" ~ Country, not needed in time series format,
#  needed for data file: covid_19_data.csv
corona %<>% mutate(    
  `Province/State` = case_when(
    is.na(`Province/State`) ~ Country,
    TRUE ~ `Province/State`)) %>% 
  arrange(Country, `Province/State`) %>% 
  group_by(count_type, Country, `Province/State`)

# provide last line of each `Province/State`
# slice(corona, n())


```


```{r}


# provide latest cases for each province, country and worldwide
# note: last(count) does not work for provinces w7o latest data => use max()
corona_province_last <- corona %>% 
  summarise(Count = max(count, na.rm = TRUE))

# for sum( , na.rm = TRUE) removing missing values is important,
# otherwise one region with NA (no cases) resulty in NA worldwide
corona_country_last <- corona_province_last %>% 
  summarise(Count = sum(Count, na.rm = TRUE))

corona_world_last <- corona_country_last %>% 
  summarise(Count = sum(Count, na.rm = TRUE))

```

```{r sum province data for country and country for world}


corona_country_sum <- corona %>% 
  group_by(count_type, Date, Country)  %>% 
  summarise(Count = sum(count))
corona_country_sum %<>% 
  group_by(count_type, Country) %>% 
  mutate(Daily_Change = c(NA, diff(Count)))

corona_world_sum <- corona %>% 
  group_by(count_type, Date)  %>% 
  summarise(Count = sum(count))
corona_world_sum %<>% 
  group_by(count_type) %>% 
  mutate(Daily_Change = c(NA, diff(Count)))

```


```{r}
ggts_counts_facet <- function(data, x = Date, y = Count) {
  col_scheme <- "Set1" # "RdYlGn" #"YlOrRd" #"Oranges" # "YlGnBu" # 
  ggplot(data, aes_(substitute(x), substitute(y), col = substitute(count_type))) +
    facet_wrap(vars(count_type), ncol = 1, scales = "free",
               strip.position = "left") +
    geom_point(size = 1.5, na.rm = TRUE) +
    geom_line(size = 1, na.rm = TRUE) +  
    theme(legend.position = "none")  +
    labs(y = "") + 
    scale_x_date(date_labels = "%b %d", date_breaks = "14 days") +
    # scale_colour_distiller(palette = col_scheme, direction = 1) +
    scale_colour_brewer(palette = col_scheme, direction = 1) +
    # scale_color_discrete(c("blue",  "green", "red")) +
    ggtitle("Confirmed, Death and Recovered Cases")
}

```


# World Map - Total Confirmed for each Country


```{r World Map}
#plot T-confirmed in world
highchart() %>%
  hc_add_series_map(worldgeojson, 
                    filter(corona_country_last, count_type == "Confirmed"), 
                    value = 'Count', 
                    joinBy = c('name', 'Country'))  %>% 
  #hc_colors(c("darkorange", "darkgray")) %>% 
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = "Countries with nCov exposure") %>% 
  hc_subtitle(text = 'with Total Confirmed - Actual Figures')
```

# Top Ten Bar Chart - Total Confirmed and Total Deaths

```{r Top Ten Bar Chart}
# Visualization with bar chart

#Top 10 countries of confirmed 
filter(corona_country_last, count_type == "Confirmed") %>%
  arrange(desc(Count)) %>% 
  head(10) %>%
  hchart("bar",hcaes(x = Country,  y = Count)) %>%
  hc_title(text = "Total Confirmed - Top Ten Countries") %>% 
  hc_add_theme(hc_theme_sandsignika())

title <- "Top 10 countries of death"
filter(corona_country_last, count_type == "Deaths") %>%
  arrange(desc(Count)) %>% 
  head(10) %>%
  hchart("bar",hcaes(x = Country,  y = Count, title = title)) %>%
  hc_title(text = "Total Deaths - Top Ten Countries") %>% 
  hc_add_theme(hc_theme_sandsignika())
```

# Diagram - Cumulated Cases and Daily Changes over Time

```{r  Time Series Diagram, fig.width = 10, fig.height = 6}
i <- "Worldwide"
plot_count <- ggts_counts_facet(corona_world_sum) + 
      labs(title = paste(i, "- Cumulated Cases"))
plot_daily_change <- ggts_counts_facet(corona_world_sum, y = Daily_Change) +
        labs(title = paste(i, "- Daily Changes"))
gridExtra::grid.arrange(plot_count, plot_daily_change, ncol = 2)

countries <- c("China", "Italy", "South Korea", "Germany")
for (i in countries) {
  plot_count <- 
    ggts_counts_facet(corona_country_sum %>% filter(Country == i)) +
      labs(title = paste(i, "- Cumulated Cases")) 
  plot_daily_change <- ggts_counts_facet(
    corona_country_sum %>% filter(Country == i), y = Daily_Change) +
            labs(title = paste(i, "- Daily Changes"))
gridExtra::grid.arrange(plot_count, plot_daily_change, ncol = 2)
}

```


# Backup

## Data Source

**Johns Hopkins University** data files on GitHub  
<https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series>

* Data files:  
*time_series_covid_19_confirmed.csv*,  
*time_series_covid_19_deaths.csv*  
*time_series_covid_19_recovered.csv*

The data are visualized on their excellent Dashboard  
**Johns Hopkins University Dashboard**  
<https://coronavirus.jhu.edu/map.html>

Files are also provided on  
**Novel Corona Virus 2019 Dataset - Day level information on covid-19 affected cases**  
<https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset>


Further links:

**WHO Dashboard**  
<https://experience.arcgis.com/experience/685d0ace521648f8a5beeeee1b9125cd>

**Robert Koch Institut, Germany**  
<https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Fallzahlen.html?nn=13490888>


## Code Source 

Code is based on ideas from 
<https://rpubs.com/TimoBoll/583802>
