---
title: Coronavirus SARS-CoV-2 / Covid-19 
subtitle: Time Series of Confirmed and Death Cases
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 12
    number_sections: yes
    toc: yes
    toc_depth: 3
urlcolor: blue
papersize: a4

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 7, fig.asp = 0.618)

#                      message = FALSE, fig.width = 7, fig.height = 4)

```


```{r initialization, eval = TRUE, include = FALSE}
#           fig.width=12, fig.height=6 ## old: fig.width=14.4, fig.height=7.2
################################################################################
## Plot and Analyze                                                           ##
## Corona Virus -Time Series of Confirmed, Death and Recovered Cases          ##
##                                                                            ##
##                                              Wolfgang Vollmer, March 2020  ##
##                                                                            ##
################################################################################

######  variables clean up, library() and	source() statements
# setwd("D:/Wolfgang/Programs-R/R-WeatherAnalysis")
setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis/Corona-Virus")
rm(list=ls()) # deletes all existing objects / variables !!
Prog.Start <- Sys.time()

library(tidyverse)
library(magrittr)
library(lubridate)
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(stlplus) # Enhanced Seasonal Decomposition of Time Series by Loess
library(DT)      # R interface to the JavaScript library DataTables
library(rlang)
library(gridExtra)

library(janitor)
library(highcharter)


######  Plot and util functions
# source("./../uts_TimeSeries.R")  # utility functions for time series
# source("./../ggts_TimeSeries.R") # ggplot2 functions for time series plots

theme_replace(
  plot.title = element_text(
    hjust = 0.5, face = "bold.italic", color = "darkcyan", size = 12),
  plot.subtitle = element_text(
    hjust = 0.5, face = "italic", color = "darkcyan")
)
x_axis_theme <- element_text(size = 14)
```

### Source data files

Data files are provided by **Johns Hopkins University** on GitHub  
<https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series>


```{r read data bind rows and chanfge to long format}

# source data file: covid_19_data.csv from
# <https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/>


# https://github.com/CSSEGISandData/COVID-19.git
# corona_confirmed <- read_csv(
#   "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv") %>%
#   mutate(Case_Type = "Confirmed") 
# corona_deaths <- read_csv(
#   "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv") %>%
#   mutate(Case_Type = "Deaths")
# corona_recovered <- read_csv(
#   "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv") %>%
#   mutate(Case_Type = "Recovered")

corona_confirmed <- read_csv(
  "./csse_covid_19_time_series/time_series_covid19_confirmed_global.csv") %>%
  mutate(Case_Type = "Confirmed")
corona_deaths <- read_csv(
  "./csse_covid_19_time_series/time_series_covid19_deaths_global.csv") %>%
  mutate(Case_Type = "Deaths")
# corona_recovered <- read_csv(
#   "./novel-corona-virus-2019-dataset/time_series_19-covid-recovered.csv")  %>%
#   mutate(Case_Type = "Recovered")


corona_wide <- bind_rows(corona_confirmed, corona_deaths)

corona <- corona_wide  %>%
    pivot_longer(cols = c(-`Province/State`, -`Country/Region`, -Case_Type, 
                          -Lat,  -Long ), 
                 names_to = c("Date"),
                 values_to = "Cases")

corona %<>% mutate(Date = as.Date(Date, format = "%m/%d/%y"))
corona %<>% rename(Country = `Country/Region`) 

```


```{r data clean-up}

# clean-up column names
# corona %>% janitor::clean_names()


# clean-up content of column Country
corona %<>% mutate(
  Country = case_when(
  Country == "Mainland China" ~ "China",
  Country == "Hong Kong SAR" ~ "Hong Kong",
  Country == "Iran (Islamic Republic of)" ~ "Iran",
  Country == "Macao SAR" ~ "Macao",
  Country == "occupied Palestinian territory" ~ "Palestine",
  Country == "Republic of Korea" ~ "South Korea",
  Country == "Korea, South" ~ "South Korea",
  Country == "Taipei and environs" ~ "Taiwan",
  Country == "US" ~ "United States of America",
  Country == "Viet Nam" ~ "Vietnam",
  TRUE ~ as.character(Country)))

# clean-up content of column `Province/State`
# replace NAs by Country
# replace "Bavaria"  by  Country since `Province/State` exists only until 
# end of January, as begin of February inluded in "Germany" data
# `Province/State` == "Bavaria" ~ Country, not needed in time series format,
#  needed for data file: covid_19_data.csv
corona %<>% mutate(    
  `Province/State` = case_when(
    is.na(`Province/State`) ~ Country,
    TRUE ~ `Province/State`)) %>% 
  arrange(Country, `Province/State`) %>% 
  group_by(Case_Type, Country, `Province/State`)

```


```{r provinces to country data}
# add all data from provinces to country data 
corona_country <- corona %>% 
  group_by(Country, Date, Case_Type) %>% 
  summarise(Cases = sum(Cases, na.rm = TRUE))
corona_country %<>% 
  group_by(Country, Case_Type) %>% 
  mutate(Daily_Change = c(NA, diff(Cases)))


corona_world <- corona_country %>% 
  group_by(Date, Case_Type) %>% 
  summarise(Cases = sum(Cases, na.rm = TRUE))
corona_world %<>% 
  group_by(Case_Type) %>% 
  mutate(Daily_Change = c(NA, diff(Cases)))

last_date <- max(corona_country$Date)

# corona_country %>% filter(Date == last_date)
# corona_world %>% filter(Date == last_date)


# provide latest cases for each province, Country and worldwide
# note: last(Cases) does not work for provinces w7o latest data => use max()
corona_province_last <- corona %>% 
  filter(Date == last_date)

# for sum( , na.rm = TRUE) removing missing values is important,
# otherwise one region with NA (no cases) resulty in NA worldwide
corona_country_last <- corona_country %>% 
  filter(Date == last_date)

corona_world_last <- corona_world %>% 
  filter(Date == last_date)

```

```{r test slice views, eval = FALSE}

sel_country <-  "United States of America" # "Germany"  "France" # "Japan" #
n_days <- 9
last_n_days <- last_date - n_days + 1  

View(corona %>% 
       filter(Country == sel_country & Date >= last_n_days) %>% 
       arrange(Case_Type)
       )
selection <- corona %>% filter(Country == sel_country & Date >= last_n_days)
(slice(selection, n() -4:0))

View(corona_country %>% 
       filter(Country == sel_country & Date >= last_n_days) %>% 
       arrange(Case_Type)
       )
selection <- corona_country %>% filter(Country == sel_country & Date >= last_n_days)
(slice(selection, n() -4:0))

```


```{r plot functions}
ggts_cases_facet <- function(data, x = Date, y = Cases, col = Case_Type) {
  col_scheme <- "Set1" # "RdYlGn" #"YlOrRd" #"Oranges" # "YlGnBu" # 
  ggplot(data, aes_(substitute(x), substitute(y), col = substitute(Case_Type))) +
    facet_wrap(vars(Case_Type), ncol = 1, scales = "free",
               strip.position = "left") +
    geom_point(size = 1.5, na.rm = TRUE) +
    geom_line(size = 1, na.rm = TRUE) +  
    theme(legend.position = "none")  +
    labs(y = "") + 
    scale_x_date(date_labels = "%b %d", date_breaks = "14 days") +
    # scale_colour_distiller(palette = col_scheme, direction = 1) +
    scale_colour_brewer(palette = col_scheme, direction = 1) +
    # scale_color_discrete(c("blue",  "green", "red")) +
    ggtitle("Confirmed and Death Cases")
}

# function to get grid plot with Cases trend and Daily_Change 
gg_trend_daily <- function(data, i) {
    plot_cases <- 
      ggts_cases_facet(data, y = Cases) +
        labs(title = paste(i, "- Cumulated Cases")) 
    plot_daily_change <- ggts_cases_facet(data, y = Daily_Change) +
              labs(title = paste(i, "- Daily Changes"))
  gridExtra::grid.arrange(plot_cases, plot_daily_change, ncol = 2)
  }

```


# World Map

## Total Confirmed for each Country

```{r World Map Confirmed}

world_map_plot <- function(data, i) {
  highchart() %>%
    hc_add_series_map(worldgeojson, 
                      data, 
                      value = 'Cases', 
                      joinBy = c('name', 'Country'))  %>% 
    #hc_colors(c("darkorange", "darkgray")) %>% 
    hc_colorAxis(stops = color_stops()) %>% 
    hc_title(text = "Spread of Coronavirus SARS-CoV-2") %>% 
    hc_subtitle(text = paste("Total", i, "/ Country - Actual Figures"))
}

# plot total confirmed in world
i <- "Confirmed"
world_map_plot(filter(corona_country_last, Case_Type == i), i)

```


## Total Deaths for each Country


```{r World Map Deaths}

# plot total deaths in world
i <- "Deaths"
world_map_plot(filter(corona_country_last, Case_Type == i), i)

```

# Bar Chart - Descending Order

## Total Confirmed

```{r Top Ten Bar Chart Confirmed}
# Visualization with top 10 country bar chart

# plot bar chart deaths / country
i <- "Confirmed"

bar_chart_countries <- function(data, i) {
  data %>%
    arrange(desc(Cases)) %>% 
    head(10) %>%
    hchart("bar",hcaes(x = Country,  y = Cases)) %>%
    hc_title(text = paste("Total", i, "- Descending Order")) %>% 
    hc_add_theme(hc_theme_sandsignika())
}

bar_chart_countries(filter(corona_country_last, Case_Type == i), i)


```

## Total Deaths

```{r Top Ten Bar Chart Deaths}

# plot bar chart deaths / country
i <- "Deaths"

bar_chart_countries(filter(corona_country_last, Case_Type == i), i)

```


# Trend / Daily Change


```{r}
i <- "Worldwide"
```

## `r i`

**Cumulated Cases and Daily Changes over Time - `r i`**

```{r  Time Series Diagram Worldwide, fig.width = 10, fig.height = 6}


i <- "Worldwide"
gg_trend_daily(corona_world, i)

```

```{r}
          
compare_countries <- c("Austria", "France", "Germany", "Italy", 
                       "Spain",  "United States of America")
countries <- c("China", "South Korea", compare_countries)

i <- 1

```


## Selected Countries

**Cumulated Cases / Daily Changes over Time -**

**`r countries`**

```{r  Time Series Diagram countries, fig.width = 10, fig.height = 6}

for (i in countries) {
gg_trend_daily(corona_country %>% filter(Country == i), i)
}

```


# Comparison Exponential Growth

## Comparison of Exponential Growth on log10 scale

```{r Comparion selected Country, fig.width = 10, fig.height = 5}
reg_range <- 9   # important no Cases value = 0 are llowed => log(0) = inf !
forcast_range <- 14  # 14days forecast => unclass(last_date + 1 + forcast_range -1)

corona_tsbl <- corona_country  %>% 
  as_tsibble(index = Date, key = c(Country, Case_Type))

corona_tsbl_filt <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% compare_countries & 
           Date >= last_date + 1 - reg_range &
           Case_Type != "Recovered") %>% 
  select(Country, Case_Type, Date, Cases)

# linear regression of log data

corona_tsbl_filt %>% 
  ggplot(aes(x=Date, y=log10(Cases))) +
  labs(x = "Date",
       title = "Comparison selected Countries", 
       subtitle ="Cumulated Cases with Linear Regression (log10 scale)") +
  geom_point() +
  geom_line(aes(col = Country), size = 1) +
  geom_smooth(method="lm", aes(col = Country), lty = "dashed", se=FALSE) +
  theme(legend.position = "bottom") +
  facet_wrap(vars(Case_Type), ncol = 2, scales = "free",
             strip.position = "left") -> gg_plot

labels_scale <- c(10, 100, 1000, 5000, 10000, 50000, 100000, 500000)
  
gg_plot + scale_y_continuous(
  breaks = log10(labels_scale),
  label = labels_scale) +
  labs(y = "Cases")


```

## Explanation

### Estimation speed of spread of the Coronavirus with Linear Regression

**Exponential Growth and Doubling Time $T$**

Exponential growth over time can be fitted by linear regression if the logarithms
of the case numbers is taken. Generally, exponential growth corresponds to
linearly growth over time for the log (to any base) data. 

The semi-logorithmic plot with base-10 log scale for the Y axis shows 
functions following an exponential law $y(t) = y_0 * a^{t/\tau}$
as straight lines. The time constant $\tau$ describes the time required for y to increase by one factor of $a$. 

If e.g. the confirmed or death cases are growing in $t-days$ by a factor of $10$
the doubling time $T \widehat{=} \tau$ can be calculated with $a \widehat{=} 2$ by

<p style="text-align: center;">
$T[days] = \frac {t[days] * log_{10}(2)} {log_{10}(y(t))-log_{10}(y_0)}$
</p>

with    
<p style="text-align: center;">
$log_{10}(y(t))-log_{10}(y_0) = = log_{10}(y(t))/y_0) = log_{10}(10*y_0/y_0) = 1$  
</p>

and doubling time   
<p style="text-align: center;">
$T[days] = t[days] * log_{10}(2) \approx t[days] * 0.30$.
</p>

For Spain, Italy, Germany we have had a doubling time up to $T \approx 9-12 days * 0.3 \approx 2.7 - 4 days !!$. 

The *doubling time $T$* and the *Forecast* is calculated for following
selected countries: **`r compare_countries`**

```{r plot function with linear regression}

ggts_cases_lm <- function(data, x = Date, y = Cases, col = Case_Type) {
  col_scheme <- "Set1" # "RdYlGn" #"YlOrRd" #"Oranges" # "YlGnBu" # 
  ggplot(data, aes_(substitute(x), substitute(y), col = substitute(Case_Type))) +
    facet_wrap(vars(Case_Type), ncol = 1, scales = "free",
               strip.position = "left") +
    geom_line(size = 1, na.rm = TRUE) +  
    geom_smooth(data = data %>% filter(Date >= last_date + 1 - reg_range), 
                method="lm", 
                col = "black", size = 1, se=FALSE) +
    geom_point(size = 1.5, na.rm = TRUE) +
    theme(legend.position = "none")  +
    labs(y = "") + 
    scale_x_date(date_labels = "%b %d", date_breaks = "14 days") +
    # scale_colour_distiller(palette = col_scheme, direction = 1) +
    scale_colour_brewer(palette = col_scheme, direction = 1) +
    # scale_color_discrete(c("blue",  "green", "red")) +
    ggtitle("Confirmed and Death Cases")
}

```


### Germany - Example plot with ~linear slope on a log10 scale

```{r  Plot with Linear Regression, fig.width = 10, fig.height = 6}

sel_country <- "Germany"

for (i in sel_country) { 
  plot_cases <- 
    ggts_cases_facet(corona_country %>% 
                        filter(Country == i & Case_Type != "Recovered")) +
      labs(title = paste(i, "- Cumulated Cases")) +
    scale_y_log10()
  plot_daily_change <- ggts_cases_lm(
    corona_country %>% 
      filter(Country == i & Case_Type != "Recovered" & Date >= last_date + 1 - 28)) +
            labs(title = paste("w/ linear regression of past", reg_range, "days"))  +
    scale_y_log10() +
    scale_x_date(date_labels = "%b %d", date_breaks = "7 days")
gridExtra::grid.arrange(plot_cases, plot_daily_change, ncol = 2)
}

```


### `r sel_country` - Trend with Forecast on a linear scale



```{r example, fig.width = 10, fig.height = 5}

fit_corona <- corona_tsbl_filt %>% ungroup() %>% 
  group_by(Case_Type) %>% 
  model(TSLM(log(Cases) ~ Date)) 
fc_corona <- fit_corona %>% fabletools::forecast(h = "14 days") 

fc_corona %>%  
  filter(Country == sel_country) %>% 
  autoplot(corona_tsbl_filt) +
  ggtitle(paste(sel_country, "- Cases and 14-days Forecast on linear y-scale"),
            subtitle = "Example shows the extreme fast increase on linear scale") +
  labs(x = "Days") +
  # scale_y_log10() +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free",
             strip.position = "left")

```


### Forecast Plot - next `r forcast_range` days {data-height=150}

The plot shows the extreme forecast increase in case of unchecked exponentiell 
growth. The dark shaded regions show 80% rsp. 95% prediction intervals. These prediction intervals are displaying the uncertainty in forecasts based on the 
linear regression over the past `r reg_range` days.


# Doubling Time and Forecast

The forecasted cases for the next 14 days are calculated 'only' from the
linear regression of the logarithmic data and are not considering any effects of 
measures in place. In addition data inaccuracies are not taken into account, 
especially relevant for the confirmed cases. 

Therefore the 14 days forecast is only an indication for the direction of an 
unchecked exponentiell growth.

```{r calculate forecast with beta_0 and beta_1, eval = TRUE}

coefs <- tidy(fit_corona) %>% 
  group_by(Country, Case_Type) %>% 
  mutate(T_doubling = (log(2) /estimate))

forecast_country <- coefs %>% 
  pivot_wider(id_cols = c(Country, Case_Type), 
              names_from = term, 
              values_from = estimate) %>% 
  rename(beta_0 = `(Intercept)`,
         beta_1 = Date) %>% 
  mutate(T_doubling = (log(2) / beta_1),
         FC_next_day = exp(beta_0 + unclass(last_date + 1) * beta_1),
         FC_14days = exp(beta_0 + unclass(last_date + 1 + forcast_range-1) 
                              * beta_1)) 

forecast_table <- forecast_country %>%
  filter(Case_Type != "Recovered") %>% 
  group_by(Country, Case_Type) %>%
  summarise(T_doubling = sum(T_doubling, na.rm = TRUE),
            FC_next_day = sum(FC_next_day, na.rm = TRUE),
            FC_14days = sum(FC_14days, na.rm = TRUE)) %>% 
  arrange(Case_Type)


countries_last_day <- corona_country %>% filter(Country %in% compare_countries & 
                                         Date == last_date &
                                         Case_Type != "Recovered") %>%
  rename(last_day = Cases) %>% 
  arrange(Case_Type) %>% 
  select(Country, Case_Type, last_day)

forecast_table <- left_join(forecast_table, countries_last_day) %>% 
  select(Country, Case_Type, T_doubling, last_day, 
         FC_next_day, FC_14days)
  

knitr::kable(forecast_table,  digits = c(0, 0, 1, 0, 0),
             caption = 
             "Forecast (FC) with linear regression: Doubling Time (days), 
             Forecasted cases tomorrow and  Forecasted cases in 14 days")

# check
# corona %>%  filter(Country %in% compare_countries, Date >= last_date + 1 - 1) 

```

### Check of Forecast Accuracy 

The forecast accuracy is checked by using the forecast method for the nine days
before the past three days (*training data*). Subsequent forecasting of the 
past three days enables comparison with the real data of these days (*test data*).

The comparison is also an early indicator if the *exponential growth* is declining.

### Germany - Forecast Accuracy for past three days

```{r forecasts accuracy, fig.width = 10, fig.height = 5}
# Note: as long as zero deaths are in this time range log(0) will go to infinity,
# therefore still no linear regression possible.

training_data_start <- reg_range + 3  # minus 12 days
training_data_end   <- reg_range - 6  # minus  3 days
# total = training + test_data => 
#    Date >= last_date + 1 - training_data_start = last 3 days

corona_tsbl_train <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% sel_country & 
           Date >= last_date + 1 - training_data_start &
           Date < last_date + 1 - training_data_end &
           Case_Type != "Recovered") %>% 
  select(Country,  Case_Type, Date, Cases)

corona_tsbl_total <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% sel_country  & 
           Date >= last_date + 1 -  training_data_start &
           Case_Type != "Recovered") %>% 
  select(Country,  Case_Type, Date, Cases)

fit_corona_train <- corona_tsbl_train %>% ungroup() %>% 
  group_by(Case_Type) %>% 
  model(TSLM(log(Cases) ~ Date)) 
fc_corona_train <- fit_corona_train %>% 
  fabletools::forecast(h = "3 days") 

# forecasts accuracy
# for evaluating accuracy on forecasts to be provided:
# - a complete dataset that includes the future data and 
# - data used to train the model.

# level = prediction interval in % or NULL
fc_corona_train %>%
  filter(Country == sel_country) %>% 
  autoplot(corona_tsbl_total, level = NULL) +
  labs(x = "Year",
       title = paste(sel_country, "Forecast Accuracy"),
       subtitle = "Forecast (blue line) vs. Data (black line)") +
    facet_wrap(~ Case_Type, ncol = 2,  scales = "free",
             strip.position = "left") +
  scale_x_date(date_labels = "%b %d", date_breaks = "2 days") +
  ggtitle("Forecast Accuracy")

```

# References

### Data Source

**Johns Hopkins University** data files on GitHub  
<https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series>

* Data files as of 2020-03-24:  
*time_series_covid19_confirmed_global.csv*,  
*time_series_covid19_deaths_global*

Note: as of 2020-03-24 new data structure which does no longer provide the 
recovered cases.

* Data files up to 2020-03-23:  
*time_series_covid_19_confirmed.csv*,  
*time_series_covid_19_deaths.csv*  
*time_series_covid_19_recovered.csv*

The data are visualized on their excellent Dashboard  
**Johns Hopkins University Dashboard**  
<https://coronavirus.jhu.edu/map.html>

Files are also provided on  
**Novel Corona Virus 2019 Dataset - Day level information on covid-19 affected cases**  
<https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset>


### Further links

**WHO Dashboard**  
<https://experience.arcgis.com/experience/685d0ace521648f8a5beeeee1b9125cd>

**Robert Koch Institut, Germany**  
<https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Fallzahlen.html?nn=13490888>

**Wikipedia - Exponential Growth**
<https://en.wikipedia.org/wiki/Exponential_growth>

### Code Source 

Code is based on ideas from 
<https://rpubs.com/TimoBoll/583802>
