---
title: Coronavirus Covid-19 / SARS-CoV-2 
subtitle: Time Series of Confirmed, Death and Recovered Cases
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 12
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
  odt_document:
urlcolor: blue
papersize: a4
params:
  city: "Basel"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 7, fig.height = 4)

```


```{r initialization, eval = TRUE, include = FALSE}
#           fig.width=12, fig.height=6 ## old: fig.width=14.4, fig.height=7.2
################################################################################
## Plot and Analyze                                                           ##
## Corona Virus -Time Series of Confirmed, Death and Recovered Cases          ##
##                                                                            ##
##                                              Wolfgang Vollmer, March 2020  ##
##                                                                            ##
################################################################################

######  variables clean up, library() and	source() statements
# setwd("D:/Wolfgang/Programs-R/R-WeatherAnalysis")
setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis/Corona-Virus")
rm(list=ls()) # deletes all existing objects / variables !!
Prog.Start <- Sys.time()

suppressMessages(library(tidyverse))
suppressMessages(library(magrittr))
library(lubridate)
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(stlplus) # Enhanced Seasonal Decomposition of Time Series by Loess
library(DT)      # R interface to the JavaScript library DataTables
library(rlang)
library(gridExtra)

library(janitor)
library(highcharter)


######  Plot and util functions
# source("./../uts_TimeSeries.R")  # utility functions for time series
# source("./../ggts_TimeSeries.R") # ggplot2 functions for time series plots

```

# Source data file: covid_19_data.csv

Downloaded from <https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/>


```{r read data and clean-up the data}

corona <- read_csv("./novel-corona-virus-2019-dataset/covid_19_data.csv")
# source data file: covid_19_data.csv from
# <https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/>

# clean-up column names
corona %<>% janitor::clean_names()

# corona %<>% mutate(obs_date = as.POSIXct(strptime(observation_date, "%m/%d/%Y")))
corona %<>% mutate(obs_date = as.Date(observation_date, format = "%m/%d/%y"))

# clean-up content of column country_region
corona %<>% mutate(
  country_region = case_when(
  country_region == "Mainland China" ~ "China",
  country_region == "Hong Kong SAR" ~ "Hong Kong",
  country_region == "Iran (Islamic Republic of)" ~ "Iran",
  country_region == "Macao SAR" ~ "Macao",
  country_region == "occupied Palestinian territory" ~ "Palestine",
  country_region == "Republic of Korea" ~ "South Korea",
  country_region == "Taipei and environs" ~ "Taiwan",
  country_region == "US" ~ "United States of America",
  country_region == "Viet Nam" ~ "Vietnam",
  TRUE ~ as.character(country_region)))

# clean-up content of column province_state
# replace NAs by country_region
# replace "Bavaria"  by  country_region since province_state exists only until 
# end of January, as begin of February inluded in "Germany" data 
corona %<>% mutate(    
  province_state = case_when(
    is.na(province_state) ~ country_region,
    province_state == "Bavaria" ~ country_region,
    TRUE ~ province_state)) %>% 
  arrange(country_region, province_state) %>% 
  group_by(country_region, province_state)

# provide last line of each province_state
# slice(corona, n())

```

```{r}

# provide latest cases for each country - summarized over all
corona_country_last <- corona %>% 
  summarise(last_prov_confirmed = last(confirmed),
            last_prov_deaths = last(deaths),
            last_prov_recovered = last(recovered)) %>% 
  summarise(sum_confirmed = sum(last_prov_confirmed),
            sum_deaths = sum(last_prov_deaths),
            sum_recovered = sum(last_prov_recovered))

corona_country_time_series <- corona %>% 
  group_by(obs_date, country_region)  %>% 
  summarise(Confirmed = sum(confirmed),
            Deaths = sum(deaths),
            Recovered = sum(recovered)) %>% 
    pivot_longer(cols = c(Confirmed, Deaths, Recovered),
                 names_to = c("count_type"),
                 values_to = "count")
# %>% ungroup()

corona_world_time_series  <- corona_country_time_series %>% 
  group_by(count_type, obs_date) %>% 
    summarise(count = sum(count))
  
  
  
```


```{r}
ggts_counts_facet <- function(data, obs_date, count, count_type) {
  ggplot(data, aes(obs_date, count, col = count_type)) +
    facet_wrap( ~ count_type, ncol = 1, scales = "free", strip.position = "left") +
    geom_point(na.rm = TRUE,  size = 1) +
    geom_line(size = 1) +  # TRUE rows w/ NA are silently removed, no warning
    theme(legend.position = "none")  +
    labs(x = "Date", y = "") + 
    scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
    ggtitle("Time Series of Confirmed, Death and Recovered Cases",
            subtitle = "Country")
}
```

```{r}
plot_world <- ggts_counts_facet(corona_world_time_series)
plot_world + labs(subtitle = "Worldwide")

plot_china <- ggts_counts_facet(
  corona_country_time_series %>% filter(country_region == "China"))
plot_china  + labs(subtitle = "China")

plot_germany <- ggts_counts_facet(
  corona_country_time_series %>% filter(country_region == "Germany"))
plot_germany + labs(subtitle = "Germany")

plot_germany <- ggts_counts_facet(
  corona_country_time_series %>% filter(country_region == "Italy"))
plot_germany + labs(subtitle = "Italy")


```


```{r orig, eval = FALSE}
#Count sum of confirmed/recover/rate of recover

# note: this version provides erroneously the cumulated sum of the data for 
# each country_region since filter for last observation date3 is missing

corona %>% 
  janitor::clean_names() %>% 
  filter(!country_region %in% 'Others') %>% 
  group_by(country_region) %>% 
  summarise(total_confirmed = sum(confirmed),
            total_recovered = sum(recovered),
            total_death = sum(deaths)) %>% 
  filter(total_confirmed > 0) %>% 
  mutate(log_total_confirmed = log(total_confirmed)) -> countries_confirmed
head(countries_confirmed)

```


# Visualization with highchart map


```{r}
#plot T-confirmed in world
highchart() %>%
  hc_add_series_map(worldgeojson, corona_country_last, value = 'sum_confirmed', joinBy = c('name','country_region'))  %>% 
  #hc_colors(c("darkorange", "darkgray")) %>% 
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = "Countries with nCov exposure") %>% 
  hc_subtitle(text = 'with Total Confirmed - Actual Figures')
```

# Visualization with bar chart

```{r}
# Visualization of `r title_rmd`

#Top 10 country of confirmed 
corona_country_last %>%
  arrange(desc(sum_confirmed)) %>% 
  head(10) %>%
  hchart("bar",hcaes(x = country_region,  y = sum_confirmed)) %>%
  hc_add_theme(hc_theme_sandsignika())
```


```{r eval = FALSE}
#plot T-confirmed in world
highchart() %>%
  hc_add_series_map(worldgeojson, countries_confirmed, value = 'total_confirmed', joinBy = c('name','country_region'))  %>% 
  #hc_colors(c("darkorange", "darkgray")) %>% 
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = "Countries with nCov exposure") %>% 
  hc_subtitle(text = 'with Total Confirmed - Actual Figures')
```


```{r eval = FALSE}
# Visualization of `r title_rmd`

#Top 10 country of confirmed 
countries_confirmed %>%
  group_by(country_region) %>%
  arrange(desc(total_confirmed)) %>% 
  head(10) %>%
  hchart("bar",hcaes(x = country_region,  y =total_confirmed)) %>%
  hc_add_theme(hc_theme_sandsignika())
```


```{r eval = FALSE}
corona %>%
  group_by(ObservationDate) %>%
  summarise(total_confirmed = sum(Confirmed),
            total_recovered = sum(Recovered),
            total_deaths = sum(Deaths))->Date_confirmed
Date_confirmed$Date <- as.Date(Date_confirmed$ObservationDate, format = "%m/%d/%y")
head(Date_confirmed)
```


```{r eval = FALSE}
#confirmed

C <- ggplot(Date_confirmed, aes(Date, total_confirmed))+
  geom_line(size = 1, alpha = 0.8,color = "#FC4E07")+
  geom_point(shape=21, color="black", fill="blue", size=2) +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days")
    
#revovered
R <- ggplot(Date_confirmed, aes(Date, total_recovered))+
  geom_line(size = 1, alpha = 0.8,color = "#FC4E07")+
  geom_point(shape=21, color="black", fill="blue", size=2) +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days")

#deaths
D <- ggplot(Date_confirmed, aes(Date, total_deaths))+
  geom_line(size = 1, alpha = 0.8,color = "#FC4E07")+
  geom_point(shape=21, color="black", fill="blue", size=2) +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days")

grid.arrange(C, R, D,ncol=1)  
```


# Backup

## Data Source 

**Novel Corona Virus 2019 Dataset - Day level information on covid-19 affected cases**

<https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset/datav>

* Data file: covid_19_data.csv
* Final Source: **Johns Hopkins University** for making the data available for educational and academic research purposes
<https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_daily_reports/>

Note: The difference between data of daily reports rsp. time_series is only 
pivot long rsp. wide format for the daily data.  
All data are from the daily case report.

**Johns Hopkins University Dashboard**
<https://coronavirus.jhu.edu/map.html>

**WHO Dashboard**
<https://experience.arcgis.com/experience/685d0ace521648f8a5beeeee1b9125cd>

**Robert Koch Institut, Germany**
<https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Fallzahlen.html?nn=13490888>


## Code Source 

Code is based on ideas from 
<https://rpubs.com/TimoBoll/583802>