---
title: "Age group dependent 7-Day Incidences for Germany"
subtitle: "based on weekly data published by RKI"
author: "Wolfgang Vollmer"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: FALSE
  pdf_document:
    fig_caption: TRUE
    number_sections: TRUE
    toc: FALSE
    toc_depth: 3
urlcolor: blue
papersize: a4
---

```{r include = FALSE}
knitr::opts_chunk$set(eval= TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 8, fig.asp = 0.618)

knitr::opts_chunk$set(comment = "#>",  collapse = TRUE)

setwd(dirname(rstudioapi::getSourceEditorContext()$path))
```

```{r include = FALSE}

library(tidyverse)
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(readxl)
library(gghighlight)
library(RColorBrewer) 
library(ggiraph) # <-- interaction to ggplot2 https://davidgohel.github.io/ggiraph
library(htmlwidgets)
library(patchwork)

# data downloaded from RKI
file <- "C:/Users/Wolfgang/Downloads/Altersverteilung.xlsx"

theme_replace(
  plot.title = element_text(
    hjust = 0.5, face = "bold.italic", color = "darkcyan", size = 12),
  plot.subtitle = element_text(
    hjust = 0.5, face = "italic", color = "darkcyan")
)
x_axis_theme <- element_text(size = 14)

# ggplot titles
title <- "Germany - Age group dependent 7-Day Incidences"
subtitle <- "- weekly new infections / 100,000 inhabitants"



```

Data source:
the age group specific incidence data is provided weekly, every Tuesday evening [RKI - COVID-19-Fälle nach Altersgruppe und Meldewoche](https://www.rki.de/DE/Content/InfAZ/N/Neuartiges_Coronavirus/Daten/Altersverteilung.html).

Attention: unfortunately RKI is changing more or less frequently table sheet name and even rownames of column "Altersgruppe".

```{r data reading and select highlighting}
age_data <- read_xlsx(file, sheet = "7-Tages-Inzidenz", col_names = TRUE)

age_data_long <- age_data %>%
  pivot_longer(cols = -Altersgruppe,
               names_to = "Calendar_Week",
               values_to = "Age_Incidence") %>% 
  separate(Calendar_Week, sep = "_", into = c("Year", "CW")) %>% 
  mutate(Year = as.integer(Year),
         CW = as.integer(CW)) %>% 
  rename(Age_group = Altersgruppe) %>% 
  unite("Year_Week", Year, CW, sep = " W") 

age_data_long$Year_Week <- yearweek(age_data_long$Year_Week)

age_data_long$Age_group <- as.factor(age_data_long$Age_group) %>% 
  recode("Gesamt" = "all") %>% 
  fct_relevel("0 - 4", "5 - 9", "10 - 14" )

age_data_long <- age_data_long %>% 
  filter(Year_Week >= yearweek("2020 W40"))

age_data_last <- age_data_long %>% 
  filter(Year_Week >= yearweek("2021 W05"))

# define age groups to be highlighted 
highlight <- c("all",  "90+", "85 - 89", "80 - 84", "75 - 79", "70 - 74", 
               "65 - 69", "60 - 64",  "10 - 14", "5 - 9", "0 - 4")
# highlight %in% unique(age_data_long$Age_group)

```

## Line and Heatmap plot from the beginning of October (CW40) 2020

```{r data long reordered}

# Lasagna plot!

# Sort with respect to value of last calendar week
no_weeks <- length(unique(age_data_long$Year_Week))
age_data_reord <- age_data_long %>% 
  mutate(Age_group = fct_reorder(.f = Age_group, .x = Age_Incidence, 
                                 # .fun = mean, na.rm = TRUE))
                                 .fun = function(x) max(x[no_weeks])))
```

```{r produce spaghetti plot long reordered}

#produce spaghetti plot:

gg_spaghetti <- ggplot(age_data_long,
                       aes(x = Year_Week, y = Age_Incidence, col = Age_group)) +
  # geom_point() +
  geom_line() +
  ylab("weekly new infections per 100,000 inhabitants") +
  guides(colour = guide_legend(reverse = TRUE)) +
  ggtitle(title, subtitle = subtitle)

gg_spaghetti <- gg_spaghetti + 
  geom_line_interactive(aes(tooltip = Age_group, data_id = Age_group)) +
    geom_point_interactive(
    aes(tooltip = Age_Incidence, data_id = Age_group), #<-- !!
    alpha = .5)

# ggiraph::girafe(code = print(gg_spaghetti), #<-- !!
#                 width_svg = 8, height_svg = 5,
#                 # define CSS properties of interactive elements
#                 options = list( 
#                   # alpha transparency =.1 for non-highlighted ("inverse hover")
#                   opts_hover_inv(css = "opacity:0.3;"),  
#                   # set line/border color and thickness for highlighted
#                   opts_hover(css = "stroke:black;stroke-width:3;") ))

```

```{r produce lasagna plot long reordered}

#produce lasagna plot:
gg_lasagna <- ggplot(age_data_long, 
                     aes(x = Year_Week, y = Age_group, 
                         fill = Age_Incidence)) + 
  geom_raster() + # geom_tile() ~ = geom_raster
  scale_fill_viridis_c("Incidence", option = "C", direction = -1) + 
  # scale_fill_gradientn(, colors = brewer.pal(6, "YlOrBr")) +
  labs(x = "Calendar week", y = "Age group / years", 
       caption = 
         "(Age groups ordered by age plus incidence data of all)") + 
  ggtitle(title, subtitle = subtitle)
# gg_lasagna

gg_lasagna <- gg_lasagna +
  geom_tile_interactive(    #<-- !!
    aes(tooltip = Age_Incidence, data_id = Age_group), 
    colour = 'transparent')
#<-- !! define grouping variable tooltip (text/value) and for hover-highlight

# ggiraph::girafe(ggobj = gg_lasagna)

# ggiraph::girafe(code = print(gg_lasagna), #<-- !!
#                 width_svg = 8, height_svg = 5,
#                 # define CSS properties of interactive elements
#                 options = list( 
#                   # alpha transparency =.1 for non-highlighted ("inverse hover")
#                   opts_hover_inv(css = "opacity:0.3;"),  
#                   # set line/border color and thickness for highlighted
#                   opts_hover(css = "stroke:black;stroke-width:2;") ))

```


```{r combined-interactive-patchwork plot long reordered}
girafe(code = print(gg_spaghetti / gg_lasagna),
       width_svg = 8, height_svg = 10,
                # define CSS properties of interactive elements
                options = list( 
                  # alpha transparency =.1 for non-highlighted ("inverse hover")
                  opts_hover_inv(css = "opacity:0.3;"),  
                  # set line/border color and thickness for highlighted
                  opts_hover(css = "stroke:black;stroke-width:2;") ))

```

## Line and Heatmap plot from the beginning of February (CW05) 2021

```{r data short reordered}

# Lasagna plot!

# Sort with respect to value of last calendar week
no_weeks <- length(unique(age_data_last$Year_Week))
age_data_reord <- age_data_last %>% 
  mutate(Age_group = fct_reorder(.f = Age_group, .x = Age_Incidence, 
                                 # .fun = mean, na.rm = TRUE))
                                 .fun = function(x) max(x[no_weeks])))
```

```{r produce spaghetti plot short reordered}

#produce spaghetti plot:

gg_spaghetti <- ggplot(age_data_reord,
                       aes(x = Year_Week, y = Age_Incidence, col = Age_group)) +
  # geom_point() +
  geom_line() +
  ylab("weekly new infections per 100,000 inhabitants") +
  ggtitle(title, subtitle = subtitle)

gg_spaghetti <- gg_spaghetti + 
  geom_line_interactive(aes(tooltip = Age_group, data_id = Age_group)) +
    geom_point_interactive(
    aes(tooltip = Age_Incidence, data_id = Age_group), #<-- !!
    alpha = .5)

# ggiraph::girafe(ggobj = gg_spaghetti)
```

```{r produce lasagna plot short reordered}

#produce lasagna plot:
gg_lasagna <- ggplot(age_data_reord, 
                     aes(x = Year_Week, y = Age_group, 
                         fill = Age_Incidence)) + 
  geom_raster() + # geom_tile() ~ = geom_raster
  scale_fill_viridis_c("Incidence", option = "C", direction = -1) + 
  # scale_fill_gradientn(, colors = brewer.pal(6, "YlOrBr")) +
  labs(x = "Calendar week", y = "Age group / years", 
       caption = 
         "(Age groups ordered by descending incidence values of last week)") +
  ggtitle(title, subtitle = subtitle) 
# gg_lasagna

gg_lasagna <- gg_lasagna +
  geom_tile_interactive(    #<-- !!
    aes(tooltip = Age_Incidence, data_id = Age_group), 
    colour = 'transparent')
#<-- !! define grouping variable tooltip (text/value) and for hover-highlight

# ggiraph::girafe(ggobj = gg_lasagna)

```


```{r combined-interactive-patchwork plot short reordered}
girafe(code = print(gg_spaghetti / gg_lasagna),
       width_svg = 8, height_svg = 10,
                # define CSS properties of interactive elements
                options = list( 
                  # alpha transparency =.1 for non-highlighted ("inverse hover")
                  opts_hover_inv(css = "opacity:0.3;"),  
                  # set line/border color and thickness for highlighted
                  opts_hover(css = "stroke:black;stroke-width:2;") ))

```




