---
title: "Coronavirus SARS-CoV-2 / Covid-19"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
link-citations: yes
cite-color: "green"
bibliography: References_Corona.bib
output:
   flexdashboard::flex_dashboard:
   vertical_layout: scroll
params:
  device: "desktop"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 7, fig.asp = 0.618)

# csl: "./../Bibliography/bluebook-law-review.csl"
# csl: bluebook-law-review.csl
# csl: harvard1.csl
# csl: ieee.csl

#===============================================================================
#
#  Corona Virus -
#        Analyze Time Series of Confirmed and Death and Cases
#
#                                           Wolfgang Vollmer,  October 30, 2020
#                                               First Version:   March 12, 2020
#
#  URL:
#  https://github.com/WoVollmer/R-TimesSeriesAnalysis/tree/master/Corona-Virus
#===============================================================================
library(flexdashboard)
library(tidyverse)
library(magrittr)
library(lubridate)
library(slider) # for rolling analysis using window functions
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(slider) # sliding window functions - rolling averages, cumulative sums
library(DT)      # R interface to the JavaScript library DataTables
library(patchwork) # package for composing plots
          # https://www.data-imaginist.com/2019/patch-it-up-and-send-it-out/
library(gghighlight) 
# https://cran.r-project.org/web/packages/gghighlight/vignettes/gghighlight.html
# library(gridExtra)
# library(janitor)
library(plotly)
library(highcharter)
library(dygraphs)  # R interface to the dygraphs JavaScript charting library
library(checkmate) # to provide informative error messages
library(pkgTS) # TimeSeries Funcstion - own R package on GitHub
               # devtools::install_github("WoVollmer/pkgTS")

```

```{r setwd change, eval = FALSE}

# knit: changes working dir => path of *.RMD file => setwd() not needed !
# R code run: RStudio project R_TimeSeriesAnalysis needs 
# relative working dir change (eval = FALSE needed for knitr runs!!) 

setwd("./Corona-Virus/") 

```

```{r data path desktop vs. notebook, eval = TRUE}
# device <- "desktop" # "notebook"  # to define setwd() and data path
# defined as params  metadata in YAML header
device <- params$device

if (device == "desktop") {
  data_path <- 
    "D:/Wolfgang/WoDocs/GitHub/COVID-19/csse_covid_19_data/csse_covid_19_time_series"
  setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis/Corona-Virus")
} else if (device == "notebook") {
  data_path <- 
    "C:/Wolfgang/Programs-R/COVID-19/csse_covid_19_data/csse_covid_19_time_series"
  # attention: for "notebook" & "linux" TimesSer.. (Time + plural-s) is needed
  #                       (whyever) not TimeSer..
  setwd("C:/Wolfgang/Programs-R/R-TimesSeriesAnalysis/Corona-Virus")
} else if (device == "linux") {
  data_path <- 
    "/home/wolfgang-ubuntu/GitHub/COVID-19/csse_covid_19_data/csse_covid_19_time_series"
    setwd("/home/wolfgang-ubuntu/Programs-R/R-TimesSeriesAnalysis/Corona-Virus")
} else {
  stop("no valid device outlined")
}

```


```{r source function and population data}
# source function ggts_corona.R and uts_corona.R 
#                            - if not in package has to come after setwd()
# sources are part of library(pkgTS)
#     all needed function files via library(pgkTS)
#     - ggts_corona.R # ggplot2 functions for time series plots
#     - uts_corona.R  # utility functions for corona analysis

```



```{r initialization, eval = TRUE, include = FALSE}

# variables clean up, library() and	source() statements ----------------------
# deletes params$device => not allowed
# rm(list=ls()) # deletes all existing objects / variables !!

# Prog.Start <- Sys.time()

theme_replace(
  plot.title = element_text(
    hjust = 0.5, face = "bold.italic", color = "darkcyan", size = 12),
  plot.subtitle = element_text(
    hjust = 0.5, face = "italic", color = "darkcyan")
)
x_axis_theme <- element_text(size = 14)
```


```{r settings}

countries_eu <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus",
                  "Czech Republic", "Denmark", "Estonia", "Finland", "France",
                  "Germany", "Greece", "Hungary", "Ireland", "Italy",
                  "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
                  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia",
                  "Spain", "Sweden")

compare_countries <- c("Austria", "France", "Germany", "Italy", "India",
                       "Japan", "Spain", "EU", 
                        "United Kingdom", "United States of America")
countries <- c("China", compare_countries)
countries_exp_growth <- c("World", "China", "Austria", "France", "Germany", "Italy", 
                       "India", "Spain", "EU",  "United States of America")
country_select <- "Germany"

span <- 7 #  for rolling mean over 7 days
q <- floor(span/2) # since 2q +1 = span !

range_reg_short <- 7 # days; at least one week is necessary to get rid of weekly 
    # fluctuations, important - no case values = 0 are allowed => log(0) = inf !
range_reg_long <- 21 # days; 2 or 3 weeks if one week is to short
                           # e.g. confidence levels to large !!
# whole week is necessary to get rid of weekly fluctuations
range_forcast <- 14  # 14days forecast => unclass(last_date + 1 + range_forcast -1)

range_cfr <- 14  # 14days cfr past period length
lag_n <- 12  # average lag for confirmed -> death - for all countries

```

```{r read hopkins raw data and process}

source("./Corona_raw_data.R")
corona_data <- read_and_process_raw_data(data_path, eu = countries_eu)

```

```{r add country population numbers and rolling mean to data}

corona_data <- add_popul_rolling_mean(corona_data) 

```


```{r}

saveRDS(corona_data, "./corona_data.RDS")
# corona_data <- readRDS("corona_data.RDS")

first_date <- min(corona_data$Date)
last_date <- max(corona_data$Date)
```


```{r complete data wide format}
# wide format: 
#    Country Population Date  Confirmed Daily_Conf ... Death ...
# for easy to read datatable outputs e.g.with filter for last date

corona_data_wide <- uts_get_corona_data_wide(corona_data)

# View(corona_data_wide %>%  dplyr::filter(Date == last_date))
```


```{r calculate lagged CFR}
## `cfr_table` - calculate Case Fatality Rate (CFR in %) -----------------------
#   assuming a time lag of `lag_n`-days (~12) between Confirmed => Death
#   - CFR_past_period : of past range_cfr (=14) days
#        (=> lag by range_cfr+1 needed 
#            to subtract last_day value - day before past period value)
#   - CFR_total
#   - CFR_unlagged (unlagged total)

corona_data_wide <- corona_data_wide %>% 
  mutate(confirmed_lag = lag(Confirmed, lag_n),
         CFR_total = round(Deaths / confirmed_lag * 100, digits = 1),
         CFR_unlagged = round(Deaths / Confirmed * 100, digits = 1)) %>% 
  mutate(confirmed_lag_range_cfr = lag(confirmed_lag, range_cfr + 1),
         deaths_lag_range_cfr = lag(Deaths, range_cfr + 1),
         confirmed_past_period = confirmed_lag - confirmed_lag_range_cfr,
         deaths_past_period = Deaths - deaths_lag_range_cfr,
         CFR_past_period = round(
           deaths_past_period / confirmed_past_period * 100, digits = 1))  %>% 
  dplyr::select(!contains("_lag"), -confirmed_past_period, -deaths_past_period) 

```


```{r}

# used by world map and bar chart data
corona_data_last <- corona_data %>%
  dplyr::filter(Date == last_date)

# next 3: used bybar chart data only
corona_data_last_wide <- corona_data_wide %>%
  dplyr::filter(Date == last_date)

corona_last_abs_wide <- corona_data_last_wide %>% 
  dplyr::select(!contains(c("_100k", "CFR")), -Population)

corona_last_rel_wide <- corona_data_last_wide %>% 
  dplyr::select(Country, Date, contains("_100k"))
```


```{r map and bar chart settings}

page_2 <- "Cumulative Cases per 100,000 Inhabitants"
page_3 <- "Mean number of cases over the past seven days per 100,000 Inhabitants"
page_4 <- paste("Case Fatality Rate - CFR of past", range_cfr, 
                "days and CFR_total assuming a time lag of", lag_n, 
                "days between Confirmed => Death")

map_page_title_1 <- paste("World Map - Spread of Coronavirus SARS-CoV-2",
                        "(Data source [[@JHU2020a]](#bib), see also [[@JHU2020]](#bib))")
map_page_title_2 <- paste("World Map -",  page_2)
map_page_title_3 <- paste("World Map -",  page_3)
map_page_title_4 <- paste("World Map -",  page_4)

bar_chart_page_title_1 <- paste("Bar Chart (descending order) and Data Table")
bar_chart_page_title_2 <- paste("Bar Chart -",  page_2)
bar_chart_page_title_3 <- paste("Bar Chart -",  page_3)
bar_chart_page_title_4 <- paste("Bar Chart -",  page_4)
  
```

```{r, eval=FALSE}

# include for every flexdashboard page an own Rmd child

```


```{r Add Page_world_map, child = 'Page_world_map.Rmd'}
```


```{r Add Page_bar_chart, child = 'Page_bar_chart.Rmd'}
```


```{r Add Page_cumulative_and_daily_trend., child = 'Page_cumulative_and_daily_trend.Rmd'}
```


```{r Add Page_exp_linear_growth, child = 'Page_exp_linear_growth.Rmd'}
```


```{r Add Page_forecast, child = 'Page_forecast.Rmd'}
```


References {#bib}
=====================================

***
***
  
### Data and Code Sources {data-height=400}

**Data Source**

Data files are provided by **Johns Hopkins University** on GitHub  
<https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series>

* Data files:  
    + *time_series_covid19_confirmed_global.csv* 
    + *time_series_covid19_deaths_global*
    + *time_series_covid19_recovered_global.csv*

The data are visualized on their Dashboard  
**Johns Hopkins University Dashboard**  
<https://coronavirus.jhu.edu/map.html>


**Code Source**

GitHub link for repository of 

* *Corona_Virus_TS_Dashboard.Rmd* R Markdown source file
    + *Page_world_map.Rmd* R Markdown child file
    + *Page_bar_chart.Rmdd* R Markdown child file
    + *Page_cumulative_and_daily_trend.Rmd* R Markdown child file
    + *Page_exp_linear_growth.Rmd* R Markdown child file
* *world_population_un.RDS* R object file with UN world population data
* *References_Corona.bib* Bibtex file providing the references with the Bibtexkeys

<https://github.com/WoVollmer/R-TimesSeriesAnalysis/tree/master/Corona-Virus>


GitHub link for repository of

* R package `pkgTS` providing functions for the **T**ime **S**eries analysis.

<https://github.com/WoVollmer/pkgTS>

R installation by calling

* `install.packages("devtools")`
* `devtools::install_github("WoVollmer/pkgTS")`

The required R files are

* `ggts_corona.R` - providing the functions to create the plots
* `uts_corona.R`  - providing utility functions

### Bibliography

