---
title: "Coronavirus SARS-CoV-2 / Covid-19"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
link-citations: yes
cite-color: "green"
bibliography: References_Corona.bib
output:
   flexdashboard::flex_dashboard:
   vertical_layout: scroll
params:
  device: "desktop"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, warning = FALSE,      
                      message = FALSE, fig.width = 7, fig.asp = 0.618)

# csl: "./../Bibliography/bluebook-law-review.csl"
# csl: bluebook-law-review.csl
# csl: harvard1.csl
# csl: ieee.csl


library(flexdashboard)
library(tidyverse)
library(magrittr)
library(lubridate)
library(fpp3)  # tsibble, tsibbledata, fable, and feasts packages & some tidyverse
library(slider) # sliding window functions - rolling averages, cumulative sums
library(DT)      # R interface to the JavaScript library DataTables
library(patchwork) # package for composing plots
          # https://www.data-imaginist.com/2019/patch-it-up-and-send-it-out/
library(gghighlight) 
# https://cran.r-project.org/web/packages/gghighlight/vignettes/gghighlight.html
# library(gridExtra)
# library(janitor)
library(plotly)
library(highcharter)
library(dygraphs)  # R interface to the dygraphs JavaScript charting library

```


```{r initialization, eval = TRUE, include = FALSE}

################################################################################
##                                                                            ##
## Corona Virus -                                                             ##
##       Analyze Time Series of Confirmed, Death and Recovered Cases          ##
##                                                                            ##
##                                          Wolfgang Vollmer,   July  2, 2020 ##
##                                              First Version: March 12, 2020 ##
##                                                                            ##
## URL:                                                                       ##
## https://github.com/WoVollmer/R-TimesSeriesAnalysis/tree/master/Corona-Virus##  
################################################################################

######  variables clean up, library() and	source() statements 
# rm(list=ls()) # deletes all existing objects / variables !!
# Prog.Start <- Sys.time()


theme_replace(
  plot.title = element_text(
    hjust = 0.5, face = "bold.italic", color = "darkcyan", size = 12),
  plot.subtitle = element_text(
    hjust = 0.5, face = "italic", color = "darkcyan")
)
x_axis_theme <- element_text(size = 14)
```


```{r settings}
# device <- "desktop" # "notebook"  # to define setwd() and data path
# defined as params  metadata in YAML header
device <- params$device

compare_countries <- c("Austria", "France", "Germany", "Italy", "India",
                       "South Korea", "Spain",  "United States of America")
countries <- c("China", compare_countries)
countries_exp_growth <- c("World", "China", "Austria", "France", "Germany", "Italy", 
                       "South Korea", "Spain",  "United States of America")
country_select <- "Germany"

span <- 7 #  for rolling mean over 7 days

range_reg <- 7   # important - no Cases value = 0 are allowed => log(0) = inf !
# changed from 9 -> 7 days, 
# but whole week is necessary to get rid of weekly fluctuations
range_forcast <- 14  # 14days forecast => unclass(last_date + 1 + range_forcast -1)
range_daily_reg <- 21  # one week is to short, confidence level to large !!
range_cfr <- 14  # 14days cfr past period length
lag_n <- 11  # average confirmed lag - for all countries

```

```{r path on desktop vs. notebook, eval = TRUE}
if (device == "desktop") {
  setwd("D:/Wolfgang/Programs-R/R-TimeSeriesAnalysis/Corona-Virus")
  
  data_path <- 
    "D:/Wolfgang/WoDocs/GitHub/COVID-19/csse_covid_19_data/csse_covid_19_time_series"
  
} else if (device == "notebook") {
  # attention: here we have (whyever) TimesS.. (Time + s) instead of TimeS..
  setwd("C:/Wolfgang/Programs-R/R-TimesSeriesAnalysis/Corona-Virus")
  
  data_path <- 
    "C:/Wolfgang/Programs-R/COVID-19/csse_covid_19_data/csse_covid_19_time_series"
} else {
  stop("no valid device outlined")
}

######  Plot and util functions
######  ######  Plot and util functions
source("./ggts_corona.R") # ggplot2 functions for time series plots

# read world population for normalization per 100k inhabitants
# population_countries <- readRDS("./world_population.RDS") # data from CIA
population_countries <- readRDS("./world_population_un.RDS")

```

```{r path on desktop PC, eval = TRUE}

corona_confirmed <- 
  read_csv(paste0(data_path, "/time_series_covid19_confirmed_global.csv")) %>%
  mutate(Case_Type = "Confirmed")
corona_deaths <- 
  read_csv(paste0(data_path, "/time_series_covid19_deaths_global.csv"))  %>%
  mutate(Case_Type = "Deaths")
corona_recovered <- 
  read_csv(paste0(data_path, "/time_series_covid19_recovered_global.csv")) %>%
  mutate(Case_Type = "Recovered")
```


```{r read data bind rows and chanfge to long format}

corona_wide <- bind_rows(corona_confirmed, corona_deaths, corona_recovered)

corona <- corona_wide  %>%
    pivot_longer(cols = c(-`Province/State`, -`Country/Region`, -Case_Type, 
                          -Lat,  -Long ), 
                 names_to = c("Date"),
                 values_to = "Cases")

corona %<>% mutate(Date = as.Date(Date, format = "%m/%d/%y")) %>% 
  rename(Country = `Country/Region`) 

```


```{r data clean-up}

# clean-up column names
# corona %>% janitor::clean_names()

# clean-up names of column Country / denmark "regions" to country 
# same country names for covid-19 data and population_countries are needed
# names have to fit to names used by
# https://code.highcharts.com/mapdata/ => 
# World with Palestine areas, high resolution Demo  => Highmaps basic demo
corona %<>% mutate(
  Country = case_when(
  Country == "Burma" ~ "Myanmar",
  Country == "Mainland China" ~ "China",
  Country == "Czechia" ~ "Czech Republic",
  Country == "Denmark" & `Province/State` == "Greenland" ~ "Greenland",
  Country == "Denmark" & `Province/State` == "Faroe Islands" ~ "Faroe Islands",
  Country == "Guinea-Bissau" ~ "Guinea Bissau",
  Country == "Hong Kong SAR" ~ "Hong Kong",
  Country == "Holy See" ~ "Holy See (Vatican City)",
  Country == "Iran (Islamic Republic of)" ~ "Iran",
  Country == "Macao SAR" ~ "Macao",
  Country == "occupied Palestinian territory" ~ "Palestine",
  Country == "Republic of Korea" ~ "South Korea",
  Country == "Cote d'Ivoire" ~ "Ivory Coast",
  Country == "Congo (Brazzaville)" ~ "Republic of Congo",
  Country == "Congo (Kinshasa)" ~ "Democratic Republic of the Congo",
  Country == "Korea, South" ~ "South Korea",
  Country == "Serbia" ~ "Republic of Serbia",
  Country == "North Macedonia" ~ "Macedonia",
  Country == "Eswatini" ~ "Swaziland",
  Country == "Taiwan*" ~ "Taiwan",
  Country == "Tanzania" ~ "United Republic of Tanzania",
  Country == "US" ~ "United States of America",
  Country == "Viet Nam" ~ "Vietnam",
  TRUE ~ as.character(Country)))

# clean-up content of column `Province/State`
# replace NAs by Country
# replace "Bavaria"  by  Country since `Province/State` exists only until 
# end of January, as begin of February inluded in "Germany" data
# `Province/State` == "Bavaria" ~ Country, not needed in time series format,
#  needed for data file: covid_19_data.csv
corona %<>% mutate(    
  `Province/State` = case_when(
    is.na(`Province/State`) ~ Country,
    TRUE ~ `Province/State`)) %>% 
  arrange(Country, `Province/State`) %>% 
  group_by(Case_Type, Country, `Province/State`)

# View(corona %>% filter(Date == last_date & Case_Type == "Confirmed"))
```


```{r provinces data to country and add country world}
# add all data from provinces to country data 
corona_country <- corona %>% 
  group_by(Country, Date, Case_Type) %>% 
  summarise(Cases = sum(Cases, na.rm = TRUE))
corona_country %<>% 
  group_by(Country, Case_Type) %>% 
  mutate(Daily_Cases = c(NA, diff(Cases)))

first_date <- min(corona_country$Date)
last_date <- max(corona_country$Date)

### provide sum of all countries as country "World" data
# for sum( , na.rm = TRUE) removing missing values is important,
# otherwise one region with NA (no cases) resulty in NA worldwide
corona_country_sum <- corona_country %>% 
  group_by(Date, Case_Type) %>% 
  summarise(Cases = sum(Cases, na.rm = TRUE))
corona_country_sum %<>% 
  group_by(Case_Type) %>% 
  mutate(Daily_Cases = c(NA, diff(Cases)),
         Country = "World")

corona_country <- bind_rows(corona_country, corona_country_sum)

# without Case_Type "Recovered"  - no real value add, no consistent data
corona_country %<>% filter(Case_Type != "Recovered")
# View(corona_country %>% filter(Date == last_date & Case_Type == "Confirmed"))
```


```{r un add population per country to data}

# add world population and normalize Confirmed and Deaths per 100k inhabitants

population_countries <- population_countries %>%
  dplyr::select(Country, Population)
          
corona_population <- full_join(corona_country, population_countries) 
# corona_population <- full_join(corona_country_last, population_countries)

corona_population %<>% filter(!is.na(Date)) %>% 
  mutate(Cases_100k = round(Cases/(Population/100000), digits = 1),
         Daily_Cases_100k = round(Daily_Cases/(Population/100000), digits = 2))

```

```{r save read rds data, eval = FALSE}

# saveRDS(corona_country, "corona_country.RDS")
# saveRDS(corona_population, "corona_population.RDS")

# corona_country <- readRDS("corona_country.RDS")
# corona_population <- readRDS("corona_population.RDS")

```

```{r complete data}
corona_all_data_wide <- corona_population %>% 
  pivot_wider(names_from = Case_Type, 
              values_from = c(Cases, Daily_Cases, Cases_100k, Daily_Cases_100k)) %>% 
  rename(Confirmed = Cases_Confirmed,
         Deaths = Cases_Deaths,
         Daily_Confirmed = Daily_Cases_Confirmed,
         Daily_Deaths = Daily_Cases_Deaths,
         Conf_100k = Cases_100k_Confirmed,
         Deaths_100k = Cases_100k_Deaths,
         Daily_Conf_100k = Daily_Cases_100k_Confirmed,
         Daily_Deaths_100k = Daily_Cases_100k_Deaths) %>% 
  mutate(Conf_mp7d = round(
    stats::filter(Daily_Confirmed, 
                  filter = rep(1 / span, span), sides = 1), digits = 0),
    Deaths_mp7d = round( 
      stats::filter(Daily_Deaths, 
                    filter = rep(1 / span, span), sides = 1), digits = 0),
    Conf_100k_mp7d = round(
      stats::filter(Daily_Conf_100k, 
                    filter = rep(1 / span, span), sides = 1), digits = 2),
    Deaths_100k_mp7d = round( 
      stats::filter(Daily_Deaths_100k, 
                    filter = rep(1 / span, span), sides = 1), digits = 2)) %>% 
  dplyr::select(Country, Population, Date, 
               Confirmed, Daily_Confirmed, Conf_mp7d, Conf_100k, 
               Daily_Conf_100k, Conf_100k_mp7d, 
               Deaths, Daily_Deaths, Deaths_mp7d, Deaths_100k, 
               Daily_Deaths_100k, Deaths_100k_mp7d)

corona_country_wide <- corona_all_data_wide %>% 
  dplyr::select(!contains("_100k"), -Population)

corona_population_wide <- corona_all_data_wide %>% 
    dplyr::select(Country, Date, contains("_100k"))

# View(corona_all_data_wide %>% filter(Date == last_date))
```


```{r}
corona_country_last <- corona_country %>%  filter(Date == last_date)
corona_population_last <- corona_population %>% filter(Date == last_date)

# corona_tsbl <- corona_country  %>%
#   tsibble::as_tsibble(index = Date, key = c(Country, Case_Type))

corona_tsbl <- corona_population  %>%
  tsibble::as_tsibble(index = Date, key = c(Country, Case_Type))
```



World Map {data-navmenu="World Map"}
=====================================  

***
***

<center>**Spread of Coronavirus SARS-CoV-2**</center>

**World Map with Confirmed and Death Cases** (Data source [[@JHU2020a]](#bib), see also [[@JHU2020]](#bib))  

### Total Confirmed for each Country

```{r World Map Confirmed}
# see https://code.highcharts.com/mapdata/

# plot total confirmed in world
i <- "Confirmed"
world_map_plot(filter(corona_country_last, Case_Type == i & Country != "World"), i)

```


### Total Deaths for each Country

```{r World Map Deaths}

# plot total deaths in world
i <- "Deaths"
world_map_plot(filter(corona_country_last, Case_Type == i & Country != "World"), i)

```


World Map / Inhabitants {data-navmenu="World Map"}
=====================================

<center>**Spread of Coronavirus SARS-CoV-2**</center>

**World Map with Confirmed and Death Cases** (Data source [[@JHU2020a]](#bib), see also [[@JHU2020]](#bib))  

### Total Confirmed for each Country per 100,000 Inhabitants

```{r World Map Confirmed per 100k Inhabitants}
# see https://code.highcharts.com/mapdata/

# plot total confirmed in world
i <- "Confirmed"
world_map_plot_inh(filter(corona_population_last, Case_Type == i & Country != "World"), i)

```


### Total Deaths for each Country per 100,000 Inhabitants

```{r World Map Deaths per 100k Inhabitants}

# plot total deaths in world
i <- "Deaths"
world_map_plot_inh(filter(corona_population_last, Case_Type == i & Country != "World"), i)

```


Bar Chart {data-navmenu="Bar Chart"}
=====================================

***
***
  
**Bar Charts with descending order**

Column {data-width=400}
-------------------------------------

### Bar Chart Confirmed

```{r Top Ten Bar Chart Confirmed}

# plot bar chart deaths / country
i <- "Confirmed"
bar_chart_countries(filter(corona_country_last,
                           Country != "World" & Case_Type == i), i)

```

### Bar Chart Table Deaths

```{r Top Ten Bar Chart Deaths}

# plot bar chart deaths / country
i <- "Deaths"
bar_chart_countries(filter(corona_country_last,
                           Country != "World" & Case_Type == i), i)

# include comment
```


Column
-------------------------------------

### Table Confirmed and Deaths - Cumulated and Daily Cases

Case numbers are taken from [[@JHU2020a]](#bib). In order to compensate for the 
daily fluctuations, the mean case numbers for the past seven days (*_mp7d) were also added.

```{r print datatable in wide format}

# DT::datatable(corona_country_last %>% filter(Case_Type != "Recovered"),
#               options=list(pageLength=25))

DT::datatable(corona_country_wide %>% 
                filter(Date == last_date),
              options = list(pageLength = 50))

```


Bar Chart / Inhabitants {data-navmenu="Bar Chart"}
=====================================

***
***
  
**Bar Charts with descending order**

Column {data-width=400}
-------------------------------------

### Bar Chart Confirmed per 100,000 Inhabitants

```{r Top 20 Bar Chart Population Confirmed}

# plot bar chart Confirmed / 100k for countries and + world
i <- "Confirmed"
bar_chart_countries_pop(filter(corona_population_last, Case_Type == i), i)

```


### Bar Chart Deaths per 100,000 Inhabitants

```{r Top 20 Bar Chart Population Deaths}

# plot bar chart deaths / country
i <- "Deaths"
bar_chart_countries_pop(filter(corona_population_last, Case_Type == i), i)

```

Column
-------------------------------------

### Table Confirmed and Deaths - Cumulated and Daily Cases per 100,000 Inhabitants

Population numbers are taken from [[@UNO2020]](#bib).
In order to compensate for the daily fluctuations, the mean case numbers for the past seven days (*_mp7d) were also added.

```{r print datatable 100k in wide format}

# DT::datatable(corona_country_last %>% filter(Case_Type != "Recovered"),
#               options=list(pageLength=25))

DT::datatable(corona_population_wide %>% 
                filter(Date == last_date), 
              options = list(pageLength = 50))

```


Bar Chart - CFR {data-navmenu="Bar Chart" #cfr}
=====================================

***
***
  
**Bar Chart with descending order**

Column {data-width=400}
-------------------------------------

### Bar Chart CFR Total - Case Fatality Rate (in %)

```{r calculate lagged CFR}


corona_confirmed_lag <- corona_country_wide %>% 
  # filter(Country %in% compare_countries) %>% 
  # filter(Country == "Germany") %>% 
  select(-Daily_Confirmed, -Daily_Deaths) %>% 
  mutate(confirmed_lag = lag(Confirmed, lag_n),
         CFR_total = round(Deaths / confirmed_lag * 100, digits = 1),
         CFR_unlagged = round(Deaths / Confirmed * 100, digits = 1)) 

corona_total_CFR_lag <- corona_confirmed_lag %>% 
  filter(Date == last_date) 



corona_before <- corona_confirmed_lag %>% 
  filter(Date == last_date - range_cfr + 1) %>% 
  mutate(Deaths_before = Deaths,
         confirmed_lag_before = confirmed_lag) %>% 
  select(Country, Deaths_before, confirmed_lag_before )


cfr_table <- left_join(corona_total_CFR_lag, corona_before) %>% 
  mutate(confirmed_past_period = confirmed_lag - confirmed_lag_before,
         deaths_past_period = Deaths - Deaths_before,
         CFR_past_period = round(
           deaths_past_period / confirmed_past_period * 100, digits = 1)) %>% 
  select(Country, CFR_total, CFR_past_period, CFR_unlagged)

```


```{r Top 20 Bar Chart CFR}

# Visualization with top 10 country - CFR  bar chart
bar_chart_cfr <- function(data, i) {
  data %>%
    arrange(desc(CFR_total)) %>% 
    head(20) %>%
    hchart("bar",hcaes(x = Country,  y = CFR_total)) %>%
    hc_title(text = paste(i, "- Case Fatality Rate (in %)")) %>% 
    hc_add_theme(hc_theme_sandsignika())
}


# plot bar chart Confirmed - CFR countries and + world
i <- "CFR_total"
bar_chart_cfr(cfr_table, i)

```

### Bar Chart CFR Total (in %) - - selected countries {data-height=200}


```{r Top 20 Bar Chart CFR selected countries}

# Visualization with top 10 country - CFR  bar chart
bar_chart_cfr <- function(data, i) {
  data %>%
    arrange(desc(CFR_total)) %>% 
    head(25) %>%
    hchart("bar",hcaes(x = Country,  y = CFR_total)) %>%
    hc_title(text = paste(i, "- Case Fatality Rate (in %)")) %>% 
    hc_add_theme(hc_theme_sandsignika())
}


# plot bar chart Confirmed - CFR countries and + world
i <- "CFR_total"
bar_chart_cfr(filter(cfr_table, Country %in% compare_countries), i)

```

Column  
-------------------------------------

### Case Fatality Rate - Proportion of deaths from confirmed cases {data-height=200}

The number of confirmed cases is an early predictor of the number of deaths.
The number of today's deaths is already determined by the infections about by 
$\sim19$ days ago or respectively by the confirmed cases about by $\sim11$ days ago.
The *Confirmed infection to Death* period can be assumed to be an average of 
$11$ days [[see @RWI2020]](#bib).

However, this varies considerably depending on country-specific test rate and 
health system. In the worst health systems it is only one day for recognized 
cases, the "Confirmed" cases must be "lagged" by $\sim1$ day. 
In the best case, the time from the end of incubation period (in average $\sim5-6$ days) to death is an average $\sim19$ days. In this case, the average 
*Confirmed infection to Death* period is $\sim14$ days), the "Confirmed" cases
must be correctly "lagged" by $\sim14$ days.
For the assumed time periods see [[@RKI2020a]](#bib), [[@RKI2020b]](#bib), for Case Fatality Rate 
and Incubation Period in general see [[@Wikipedia2020a]](#bib),
[[@Wikipedia2020b]](#bib).

The simple calculation with unlagged cumulated confirmed cases divided by
cumulated deaths results in a significant underestimation of the CFR in health systems with early disease detection. 
If the number of cumulative cases is already large compared to the number of 
active cases (~ cases from the past two weeks), the "lagged" rsp. "unlagged"
values converge.

The Infection Fatality Rate (IFR) is the fatality rate of all infection, that means
detected confirmed cases and undetected cases (asymptomatic and not tested group).
This lethality is assumed to 
be country independent and only rough estimates exist (RKI: bottom of existing estimates $\sim0.56\%$). 


```{r CFR Table}

title <- paste("Case Fatality Rate in %", "- CFR_total and CFR_past_period (period of past", 
               range_cfr, "days) w/ Confirmed lagged by", lag_n, "days")
# knitr::kable(cfr_table,
#              digits = 2, caption = title)

```

### `r title` {data-height=400}

```{r}

DT::datatable(cfr_table, 
              options = list(pageLength = 10))
```

### Case Fatality Rate (in %) - selected countries {data-height=200}


```{r CFR Table selected countries}

knitr::kable(filter(cfr_table, Country %in% compare_countries),
             digits = 2)

```



Cumulated and Daily Trend {data-orientation=rows data-navmenu="Cumulated and Daily Trend" }
=====================================

***
***


```{r}

country <- "World"
fig_width_trend <- 10 # fig.width = fig_width_trend 7
fig_asp_trend <- 0.35
```


**Cumulated and Daily Cases over Time**  

Row {data-height=400}
------------------------------------

### `r country`


```{r  Time Series Diagram Worldwide, fig.width = fig_width_trend, fig.asp = fig_asp_trend}

ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- 1

```


Row {.tabset .tabset-fade}
------------------------------------

**Selected Countries**

### `r countries[i]` 

```{r  Time Series Diagram countries 1, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`

```{r  Time Series Diagram countries 2, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`


```{r  Time Series Diagram countries 3, fig.width = fig_width_trend,  fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`


```{r  Time Series Diagram countries 4, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```


### `r countries[i]`


```{r  Time Series Diagram countries 5, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`

```{r  Time Series Diagram countries 6, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`

```{r  Time Series Diagram countries 7, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`

```{r  Time Series Diagram countries 8, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### `r countries[i]`

```{r  Time Series Diagram countries 9, fig.width = fig_width_trend, fig.asp = fig_asp_trend}
country <- countries[i]
ggts_trend_daily(corona_country %>% filter(Country == country), country)

i <- i + 1
```

### Germany - Confirmed and Deaths

```{r dygraph_plot prepare, eval = TRUE}

# https://rstudio.github.io/dygraphs/gallery-series-highlighting.html
# dygraph interactive plot for time series data:
# input must be a named list or data frame, where the first element/column
# provides x-axis values and all subsequent elements/columns provide one or more
# series of y-values.
plot_dygraph_all <- function(data_xts, country_select, last_date) {
  dygraph(data_xts, 
          main = paste(country_select, "- Cumulated and Daily Cases")) %>% 
    dyAxis("y", label = "Cumulated Cases") %>%
    dyAxis("y2", label = "Daily Cases", independentTicks = TRUE) %>%
    dyLegend(width = 400) %>% 
    dySeries("Confirmed", drawPoints = TRUE, pointShape = "square") %>% 
    dySeries("Deaths", drawPoints = TRUE, pointShape = "square") %>% 
    dySeries("Daily_Confirmed", drawPoints = TRUE, pointShape = "square",
             axis = "y2") %>% 
    dySeries("Daily_Deaths", drawPoints = TRUE, pointShape = "square", 
             axis = "y2") %>% 
    dyRangeSelector(dateWindow = 
                      c(as.character(last_date - 28), as.character(last_date)))
}


plot_dygraph_daily <- function(data_xts, country_select, last_date) {
  dygraph(data_xts, 
          main = paste0(country_select, " - ",
                        span, "-day Rolling Mean of Daily Cases")) %>% 
    dyAxis("y", label = "Daily Confirmed Cases") %>%
    dyAxis("y2", label = "Daily Death Cases", independentTicks = TRUE) %>%
    dyLegend(width = 400) %>% 
    # dySeries("Daily_Confirmed", drawPoints = TRUE, pointShape = "square") %>% 
    dySeries("Daily_Conf_rol_mean", drawPoints = FALSE) %>% 
    # dySeries("Daily_Deaths", drawPoints = TRUE, pointShape = "square", 
    #          axis = "y2") %>%     
    dySeries("Daily_Deaths_rol_mean", drawPoints = FALSE, axis = "y2") %>% 
    dyRangeSelector(dateWindow = 
                      c(as.character(last_date - 2 * 28), as.character(last_date)))
}

plot_dygraph_daily_repro <- function(data_xts, country_select, last_date) {
  dygraph(data_xts, 
          main = paste0(country_select, " - ",
                        span, 
                        "-day window Reproduction Number based on Daily Confimred Cases")) %>% 
    dyAxis("y", label = "Reproduction Number W/ Confidence Interval", 
           independentTicks = TRUE) %>%
    dyAxis("y2", label = "Daily Confirmed Cases") %>%
    dyLegend(width = 400) %>%    
    dySeries(c("ci.lower","Repro_number", "ci.upper")) %>% 
    dyLimit(1, "Repro_number = 1",
            strokePattern = "dashed", color = "red") %>%
    dySeries("Daily_Confirmed", drawPoints = FALSE,
             pointShape = "square", axis = "y2") %>%
    dySeries("Daily_Conf_rol_mean", axis = "y2") %>%
    dyRangeSelector(dateWindow =
                      c(as.character(last_date - 2 * 28), as.character(last_date)))
}

data <- corona_country_wide %>% 
  filter(Country == country_select) 

# add roll_mean
span <- 7 # rolling mean over 7 days
weeks <- 6
data %<>% mutate(Daily_Conf_rol_mean = round(
                   stats::filter(Daily_Confirmed, 
                                 filter = rep(1 / span, span)), digits = 0),
                 Daily_Deaths_rol_mean = round( 
                   stats::filter(Daily_Deaths, 
                                 filter = rep(1 / span, span)), digits = 0)) 


# change tibble in xts object
index <- data$Date
data  %<>% ungroup() %>% select(-Country, -Date) 

```

```{r dygraph_plot Cumulated and Daily, eval = TRUE}

data_xts <- xts::xts(
  data %>% dplyr::select(Confirmed, Deaths, Daily_Confirmed, Daily_Deaths), 
  order.by = index)

plot_dygraph_all(data_xts, country_select, last_date)

```

Reproduction Number {data-orientation=rows data-navmenu="Cumulated and Daily Trend"}
=====================================

***
***

**`r country_select` - Rolling Mean and Reproduction Number**  

Row {data-height=400}
------------------------------------

### `r country_select` Rolling Mean of Daily Cases  

```{r dygraph_plot Daily Rolling Mean}

data_xts <- xts::xts(
  data %>% dplyr::select(Daily_Conf_rol_mean,    
                         Daily_Deaths_rol_mean), 
  order.by = index)

plot_dygraph_daily(data_xts, country_select, last_date)

```

### Rolling Mean of Daily Cases {data-width=300} 

The *7-day Rolling Mean/Moving Average* of the Daily Confirmed and Death Cases 
smooths out the short-term weekly fluctuations (weekend).

The daily confirmed cases are related to the left y-axes, the daily death cases 
are related to the right y-axes. This clearly outlines the 12 days delay relation between daily confirmed and death cases and also  the roughly the factor 
of ~1/25 (~4%).






Row {data-height=400}
------------------------------------

### `r country_select` Calculated Reproduction Number  


```{r dygraph_plot Reproduction Number Rolling Mean, eval = TRUE}


# infectivity profile w(t) w/ w(t= day_1 & day>=11)=0:
infectivity <- c((0:3)/3, 1, (5:0)/5)
names(infectivity) <- seq_along(infectivity)
infectivity <- infectivity / sum(infectivity)

width <- 7         #  => window (rolling mean one sided) in repronum() 
report.delay <- 7  #  => delay (reporting delay) in repronum() 
alpha <- 0.05      # => default conf.level in repronum() 


# add roll_mean
span <- 7 # rolling mean over 7 days
weeks <- 6

data_repronum <- corona_country_wide %>% 
  filter(Country == country_select) %>% 
  dplyr::select(Country, Date, Daily_Confirmed) %>% 
  mutate(Daily_Conf_rol_mean = round(
    stats::filter(Daily_Confirmed, 
                  filter = rep(1 / span, span)), digits = 0)) 

# repronum_output <- repronum(new.cases = data_repronum$Daily_Conf_rol_mean, 
repronum_output <- repronum(new.cases = data_repronum$Daily_Confirmed, 
                            profile = infectivity, 
                            window = width,
                            delay = report.delay)


data_repronum  <- bind_cols(data_repronum, repronum_output) %>% 
  rename(Repro_number = repronum)


# change tibble in xts object
index <- data_repronum$Date
data_repronum  %<>% ungroup() %>% select(-Country, -Date, -repronum.se) 


data_xts <- xts::xts(data_repronum,
  # data %>% dplyr::select(Daily_Conf_rol_mean, Daily_Deaths_rol_mean), 
  order.by = index)

plot_dygraph_daily_repro(data_xts, country_select, last_date)


```


### Calculated Reproduction Number {data-width=300} 

The calculation of the  **reproduction number** $R(t)$ uses 
a R function provided by [[@Hotz2020]](#bib) on GitHub. 

The (effective) **reproduction number** $R(t)$ at day $t$, 
i.e. the average number of people someone infected at time $t$ would infect 
if conditions remained the same. 

For further 
German federal states figures (based on the data provided by Robert Koch Institut) see [[@Hotz2020a]](#bib) and for
worldwide figures  (based on the data provided by Johns Hopkins University) see [[@Hotz2020b]](#bib).


For the calculation the assumption of 7-days reporting delay 
(confirmed is reported 7-days after 'real' infection) is unchanged and the same
modelled infectivity profile *w* is used.
The lower and upper confidence interval lines provide the (approximate, pointwise) 
**95% confidence interval** (only based on statistical numbers, possible changes 
in e.g. counting measures can not be considered). 

This is the reason why [[@Hotz2020]](#bib) "do not compute an average over a 
sliding window of seven days so the viewer immediately recognizes the size of such
artefacts, warning her to be overly confident in the results. In fact, these artefacts are much larger than the statistical uncertainty due to the stochastic nature of the epidemic which is reflected in the confidence intervals."

Nevertheless, here the calculation is based on the 7-days rolling mean and therefore
the figure smooths over the the weekly rhythm.


Virus Spread on log10 scale {data-navmenu="Exp Linear Growth"}
=====================================

***
***


Column {data-width=100}
-------------------------------------

### Exponential Growth Evaluation

China and South Korea slowed down exponential growth significantly at an early stage. Their lines in the diagram with the log10 scale have therefore no longer 
had a significant slope.
 
In early phases countries have a more or less unchecked exponential
growth. 
If countermeasures are effective, reduced exponential growth is reflected in a 
reduced slope of the accumulated cases again.

Column
-------------------------------------

### Virus Spread with log10 scale (since mid of Jan)

```{r}

labels_scale <- c(10, 100, 1000, 5000, 10000, 50000, 100000, 500000, 1000000, 
                  5000000, 10000000)

```


```{r plot log10scale w/o date filter, fig.width = 14, fig.asp = 0.8}

# logscale plot w/o date filter
data <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in%  c("World", "China", compare_countries) & 
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

gg_plot <- gg_logscale(data)
  
plot_1 <- gg_plot + 
  scale_y_continuous(breaks = log10(labels_scale), label = labels_scale) +
  geom_line(aes(col = Country), size = 0.5) +
  labs(title = "Cumulated Cases") +
  facet_wrap(vars(Case_Type), ncol = 1, scales = "free_y",
             strip.position = "left")

##  plot per 100,000 inhabitants
data <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in%  c("World", "China", compare_countries) & 
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Cases_100k)

gg_plot <- gg_logscale(data, x = Date, y = Cases_100k)
  
plot_2 <- gg_plot + 
  scale_y_continuous(breaks = log10(labels_scale), label = labels_scale) +
  geom_line(aes(col = Country), size = 0.5) +
  labs(title = "Cumulated Cases / 100k Inhabitants",
       y = "Cumulated Cases / 100k Inhabitants") +
  facet_wrap(vars(Case_Type), ncol = 1, scales = "free_y",
             strip.position = "left") 


plot_1 + plot_2 + 
  plot_annotation(
    title = "Virus Spread (with log10 scale) - World and selected Countries") +
  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

```



```{r plot per inhabitants with log10scale, eval = FALSE, fig.width = 7, fig.asp = 1.5}
# Column
# -------------------------------------
# 
# ### Virus Spread / Inhabitants with log10 scale (since mid of Jan)

# Virus Spread / 100,000 Inhabitantson on log10 scale (since mid of Jan)

data <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in%  c("World", "China", compare_countries) & 
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Cases_100k)

gg_plot <- gg_logscale(data, x = Date, y = Cases_100k)
  
gg_plot + 
  scale_y_continuous(breaks = log10(labels_scale), label = labels_scale) +
  geom_line(aes(col = Country), size = 0.5) +
  labs(subtitle = paste("Since", first_date),
       y = "Cumulated Cases / 100k Inhabitants",
       title = "Virus Spread / 100,000 Inhabitants (with log10 scale)") +
  facet_wrap(vars(Case_Type), ncol = 1, scales = "free_y",
             strip.position = "left") 

```

Exponential Growth {data-navmenu="Exp Linear Growth"}
=====================================

***
***

Column {data-width=400}
------------------------------------  

### Estimation spread of the Coronavirus with Linear Regression of log data

**Exponential Growth and Doubling Time $T$**

Exponential growth over time can be fitted by linear regression if the logarithms
of the case numbers is taken. Generally, exponential growth corresponds to
linearly growth over time for the log (to any base) data 
[[see also @Wikipedia2020]](#bib). 

The semi-logorithmic plot with base-10 log scale for the Y axis shows 
functions following an exponential law $y(t) = y_0 * a^{t/\tau}$
as straight lines. The time constant $\tau$ describes the time required for y to increase by one factor of $a$. 

If e.g. the confirmed or death cases are growing in $t-days$ by a factor of $10$
the doubling time $T \widehat{=} \tau$ can be calculated with $a \widehat{=} 2$ by

<p style="text-align: center;">
$T[days] = \frac {t[days] * log_{10}(2)} {log_{10}(y(t))-log_{10}(y_0)}$
</p>

with    
<p style="text-align: center;">
$log_{10}(y(t))-log_{10}(y_0) = = log_{10}(y(t))/y_0) = log_{10}(10*y_0/y_0) = 1$  
</p>

and doubling time   
<p style="text-align: center;">
$T[days] = t[days] * log_{10}(2) \approx t[days] * 0.30$.
</p>

For Spain, Italy, Germany we have had a doubling time of only $T \approx 9days * 0.3 \approx 2.7 days$ at the beginning of the pandemic!!. 

The *doubling time $T$* and the *Forecast* is calculated for following
selected countries: **`r compare_countries`** and **World** in total 
(see [Forecast / Doubling Time](#doubling)).


### `r country_select` - Trend with Forecast on a linear scale


```{r corona tsibble range_reg}


# provide tsb of range_reg/last 7 days for linear regression of log data
corona_tsbl_range_reg <- corona_tsbl %>% 
  ungroup() %>% 
  group_by(Case_Type) %>% 
  filter(Country %in% c("World", compare_countries) & 
           Date >= last_date - range_reg + 1 &
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Cases)



```

```{r corona daily tsibble range_reg}

# provide tsb of range_reg/last 7 days for linear regression of daily data
corona_range_daily_reg_tsbl <- corona_tsbl %>% 
  ungroup() %>% 
  group_by(Case_Type) %>% 
  filter(Country %in% c("World", compare_countries) & 
           Date >= last_date - range_daily_reg + 1 &
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Daily_Cases)


```


```{r example, fig.width = 10, fig.height = 5}

fit_corona <- corona_tsbl_range_reg %>% ungroup() %>% 
  group_by(Case_Type) %>% 
  model(TSLM(log(Cases) ~ Date)) 
fc_corona <- fit_corona %>% fabletools::forecast(h = "14 days") 

fc_corona %>%  
  filter(Country == country_select) %>% 
  autoplot(corona_tsbl_range_reg) +
  ggtitle(paste(country_select, "- Cases and 14-days Forecast on linear y-scale"),
            subtitle = "Example shows the extreme fast increase on linear scale") +
  labs(x = "Days") +
  # scale_y_log10() +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left")

```


### Forecast Plot - next `r range_forcast` days {data-height=150}

The plot shows the extreme forecast increase in case of unchecked exponentiell 
growth. The dark shaded regions show 80% rsp. 95% prediction intervals. These prediction intervals are displaying the uncertainty in forecasts based on the 
linear regression over the past `r range_reg` days.

Column {data-width=600}
------------------------------------ 

### Comparison Exponential Growth 


```{r plot log10scale selected Country}

gg_plot <- gg_logscale(corona_tsbl_range_reg)  +
  geom_smooth(method="lm", aes(col = Country), lty = "dashed", se=FALSE)
  
gg_plot +
  scale_y_continuous(breaks = log10(labels_scale), label = labels_scale) +
  labs(subtitle = paste("Past", range_reg, "days with Linear Regression"))

```



### Germany - Example plot with ~linear slope on a log10 scale {data-height=400}

```{r  Plot with Linear Regression}
# plot with linear regression on log scale
for (i in country_select) { 
  plot_cases <- 
    ggts_cases_facet(corona_country %>% 
                       filter(Country == i & Case_Type != "Recovered" 
                              & Cases >= 10)) +
    labs(title = paste(i, "- Cumulated Cases")) +
    scale_y_log10()
  plot_daily_cases <- 
    ggts_cases_facet(corona_country %>% 
                       filter(Country == i & Case_Type != "Recovered" & 
                                Cases >= 10 &
                                Date >= last_date - 28 + 1)) +
    labs(title = paste("w/ linear regression of past", range_reg, "days"))  +
    geom_smooth(data = corona_country %>% 
                  filter(Country == i & Case_Type != "Recovered" & 
                           Date >= last_date + 1 - range_reg), 
                method="lm", 
                col = "black", size = 1, se = FALSE) +
    scale_y_log10() +
    scale_x_date(date_labels = "%b %d", date_breaks = "7 days")
    # gridExtra::grid.arrange(plot_cases, plot_daily_cases, ncol = 2)
    print(plot_cases + plot_daily_cases) 
}

```

Compare Exp vs Linear Growth {data-navmenu="Exp Linear Growth"}
=====================================

***
***


Column {data-width=150}
-------------------------------------

### Comparison Exponential vs. Linear Growth

The charts compare the different forecasts for an exponential rsp. linear growth model. Due to the large fluctuations of the daily cases regression of three weeks is required. Otherwise the prediction levels are much too big.

The dark shaded regions are indicating the $80\%$ rsp. $95\%$  prediction 
intervals. These prediction intervals are displaying the "pure" statistical 
uncertainty in forecasts based on the regression models.

For doubling periods in the order of period of infectivity
(RKI assumption: $\sim9-10$ days, with great uncertainty, 
see [[@RKI2020b]](#bib), we no longer have exponential growth. 
The "old" infected cases are at the end of the doubling period no longer infectious (active). This results in a constant infection rate with basic reproduction number $R_t \sim 1$ or even $<1$.

Note: for case numbers of German federal states see [[@RKI2020]](#bib).


Column
-------------------------------------
### Cumulated Cases - Comparison Exponential and Linear Growth

```{r cumulated forecast exp growth, fig.asp = fig_asp_trend}

fit_cum_corona <- corona_tsbl_range_reg %>% ungroup() %>% 
  group_by(Case_Type) %>% 
  model(Exp_Growth = TSLM(log(Cases) ~ Date),
        Linear_Growth = TSLM(Cases ~ Date)) 

fc_cum_corona <- fit_cum_corona %>% fabletools::forecast(h = "7 days") 

corona_cum_range_daily_reg_tsbl <- corona_tsbl %>% 
  ungroup() %>% 
  group_by(Case_Type) %>% 
  filter(Country %in% c("World", compare_countries) & 
           Date >= last_date - range_daily_reg + 1 &
           Case_Type != "Recovered" & Cases >= 10) %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

fc_cum_corona %>%  
  filter(Country == country_select) %>% 
  autoplot(corona_cum_range_daily_reg_tsbl) +
  labs(x = "Days",
       title = paste(
         country_select, 
         "- Cumulated Cases and 7-days Forecast"),
       subtitle = 
         "Forecasts with prediction intervals based on regression of past week") +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left")

```

### Daily Cases - Comparison Exponential and Linear Growth

```{r daily forecast linear model, fig.asp = fig_asp_trend}

fit_daily_corona <- corona_range_daily_reg_tsbl %>% ungroup() %>% 
  group_by(Case_Type) %>% 
    model(Exp_Growth = TSLM(log(Daily_Cases) ~ Date),
        Linear_Growth = TSLM(Daily_Cases ~ Date)) 
  # model(TSLM(Daily_Cases ~ Date)) 
fc_daily_corona <- fit_daily_corona %>% fabletools::forecast(h = "7 days") 

fc_daily_corona %>%  
  filter(Country == country_select) %>% 
  autoplot(corona_range_daily_reg_tsbl) +
    labs(x = "Days",
       title = paste(
         country_select, 
         "- Daily Cases and 7-days Forecast"),
       subtitle = 
         "Forecasts with prediction intervals based on regression of past 3 weeks") +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left")

```



Doubling Time / Forecast {data-navmenu="Forecast" #doubling}
=====================================

***
***

Column {data-width=500}
-------------------------------------

**Doubling Time and Forecast** 
The forecasted cases for the next 14 days are calculated 'only' from the
linear regression of the logarithmic data and are not considering any effects of 
measures in place. In addition data inaccuracies are not taken into account, 
especially relevant for the confirmed cases. 

Therefore the 14 days forecast is only an indication for the direction of an 
unchecked exponentiell growth.

The doubling rate is only a good indicator at the beginning of the pandemic. If the number of confirmed cases from the past two weeks is already small compared to the total number of confirmed cases, the number of infectious people is also small compared to the total cases.

```{r calculate forecast with beta_0 and beta_1, eval = TRUE}

coefs <- tidy(fit_corona) %>% 
  group_by(Country, Case_Type) %>% 
  mutate(T_doubling = (log(2) /estimate))

forecast_country <- coefs %>% 
  pivot_wider(id_cols = c(Country, Case_Type), 
              names_from = term, 
              values_from = estimate) %>% 
  rename(beta_0 = `(Intercept)`,
         beta_1 = Date) %>% 
  mutate(T_doubling = (log(2) / beta_1),
         FC_next_day = exp(beta_0 + unclass(last_date + 1) * beta_1),
         FC_14days = exp(beta_0 + unclass(last_date + 1 + range_forcast-1) 
                              * beta_1)) 

forecast_table <- forecast_country %>%
  filter(Case_Type != "Recovered") %>% 
  group_by(Country, Case_Type) %>%
  summarise(T_doubling = sum(T_doubling, na.rm = TRUE),
            FC_next_day = sum(FC_next_day, na.rm = TRUE),
            FC_14days = sum(FC_14days, na.rm = TRUE)) %>% 
  arrange(Case_Type)

countries_last_day <- corona_country %>% 
  filter(Country %in% c(compare_countries, "World") & 
           Date == last_date &
           Case_Type != "Recovered") %>%
  rename(last_day = Cases) %>% 
  arrange(Case_Type) %>% 
  select(Country, Case_Type, last_day)

forecast_table <- left_join(forecast_table, countries_last_day) %>% 
  select(Country, Case_Type, T_doubling, last_day, 
         FC_next_day, FC_14days)

knitr::kable(forecast_table,  digits = c(0, 0, 1, 0, 0),
             format.args = list(big.mark = "'"),
             caption = 
             "Forecast (FC) with linear regression: Doubling Time (days), 
             Forecasted cases tomorrow and  Forecasted cases in 14 days")

# check
# corona %>%  filter(Country %in% compare_countries, Date >= last_date + 1 - 1) 

```

Column
-----------------------------------------------------------------------  

### Check of Forecast Accuracy {data-height=400}

The forecast accuracy is checked by using the forecast method for the nine days
before the past three days (*training data*). Subsequent forecasting of the 
past three days enables comparison with the real data of these days (*test data*).

The comparison is also an early indicator if the *exponential growth* is declining. However, possible changes in underreporting
(in particular the proportion confirmed / actually infected) requires careful interpretation.

For doubling periods in the order of infectivity (RKI assumption: $\sim9-10$
days, with great uncertainty, [[see @RKI2020b]](#bib), we no longer have exponential growth. Since the "old" infected cases are 
no longer infectious after these periods and we then have a constant infection rate with  basic reproduction number $R_t \sim 1$.

Instead, we have "only" linear growth of the cumulative *Confimred Cases* and 
the *Daily Confirmed Cases* remain more or less constant or even decrease.

However, the basic reproduction number $R_0 (\approx 2.2 - 3)$ is a product of the average number of contacts of an infectious person per day, the probability of transmission upon contacts and the average number of days infected people are infectious. 
With the current uncertainty of the average duration of the infectivity duration, $R_0$ can therefore be estimated from the doubling time only to a very limited extent.
See also [[@CMMID2020]](#bib) and [[@RKI2020c]](#bib).

### Germany - Forecast Accuracy for past three days {data-height=300}

```{r forecasts accuracy, fig.width = 10, fig.height = 5}
# Note: as long as zero deaths are in this time range log(0) will go to infinity,
# therefore still no linear regression possible.

training_data_start <- 12 #range_reg + 3  # minus 12 days
training_data_end   <- 3  #range_reg - 6  # minus  3 days
# total = training + test_data => 
#    Date >= last_date + 1 - training_data_start = last 3 days

corona_tsbl_train <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% country_select & 
           Date >= last_date + 1 - training_data_start &
           Date < last_date + 1 - training_data_end &
           Case_Type != "Recovered") %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

corona_tsbl_total <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% country_select  & 
           Date >= last_date + 1 -  training_data_start &
           Case_Type != "Recovered") %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

fit_corona_train <- corona_tsbl_train %>% ungroup() %>% 
  group_by(Case_Type) %>% 
  model(TSLM(log(Cases) ~ Date)) 
fc_corona_train <- fit_corona_train %>% 
  fabletools::forecast(h = "3 days") 

# forecasts accuracy
# for evaluating accuracy on forecasts to be provided:
# - a complete dataset that includes the future data and 
# - data used to train the model.

# level = prediction interval in % or NULL
fc_corona_train %>%
  filter(Country == country_select) %>% 
  autoplot(corona_tsbl_total) +
  labs(x = "Year",
       title = paste(country_select, "Forecast Accuracy"),
       subtitle = "Forecast (blue line) vs. Data (black line)") +
    facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left") +
  scale_x_date(date_labels = "%b %d", date_breaks = "2 days") +
  ggtitle("Forecast Accuracy")

```

Forecast {data-navmenu="Forecast"}
=====================================

***
***

Column {data-width=400}
-------------------------------------

### Forecasting with lagged Predictors

The number of confirmed cases is an early predictor of the number of deaths.
The number of today's deaths is already determined by the infections about by 
$\sim19$ days ago or respectively by the confirmed cases about by $\sim11$ days
ago [see Bar Chart - CFR](#cfr).

The country specific case fatality rate (CFR, proportion of deaths from 
confirmed cases) indicates the country specific testing rate and may depend on quality/capacity of hospitals. 

Overall a rough conclusion on the country specific **underreporting rate** 
(lack of diagnostic confirmation; proportion of all infected to confirmed cases)
is feasible if the infection fatality rate (IFR, confiremd cases plus all 
asymptomatic and undiagnosed infections) is assumed to be country independent 
and the IFR is known (bottom of existing estimates $\sim0.56\%$, assumption by 
RKI see [[@RKI2020b]](#bib).

In this case an estimation of the CFR of $0.06$ $(6\%)$ indicates an
underreporting by a by a factor of $\sim10$. 
A CFR of $0.20$ $(20\%)$ indicates an underreporting by a by a factor of $\sim30$.
This corresponds to RKI assumption of a underreporting by a factor of $11-20$ [[@RKI2020a]](#bib). Unfortunately, the IFR or lethality is still far too 
imprecise for concrete conlusions.


In the model paper RKI assumes for the

* Incubation period $\sim5-6$ days - Day of infection day until symptoms are upcoming)
* Hospitalisation $+4$ days  - Admission to the hospital (if needed) after Incubation Period)
* Average period to death $+11$ - if the patient dies, it takes an average of 
$11$ days after admission to the hospital

Depending on the country-specific test frequency (late or early tests), 
the 

*lag_days - time from receipt of the confirmed test result to death, Confirmed to Death, 
is about $11-13$ days.


Note: these methods are also used for example for advertising campaigns. The
campaign impact on sales will be some time beyond the end of the campaign, and 
sales in one month will depend on the advertising expenditure in each of the 
past few months [[see @Hyndman2020]](#bib).


Column {data-width=600}
-------------------------------------

### Daily Confirmed and Death Cases

```{r forecast for selected countries and lags}

# avg_days_to_confirmed  # Incubation period ~ 5-6 days; Hospitalisation + 4 days
# avg_days_to_death      # death     ~ after 21 days
# case_fatality_rate     # proportion of deaths / number of people diagnosed
#                                                                   (Confirmed)

lag_germany <- 12 # 13
lag_italy <-  12  # 10
lag_spain <- 10   # 11

countries_to_forecast <- c("Germany", "Italy", "Spain")

countries_fc_param <-
  tibble(Country = countries_to_forecast,
       confirmed_to_death = c(lag_germany, lag_italy , lag_spain),
       infected_to_death     = 21,
       case_fatality_rate = c(0.06, 0.25, 0.3)) %>% 
  mutate(infected_to_confirmed = infected_to_death  - confirmed_to_death)

        
```


```{r eval = TRUE}

corona_country_sel <- corona_country %>% 
  filter(Country %in% countries_to_forecast & 
           Case_Type != "Recovered" &
           Date >= last_date - 28) %>%
  ungroup() %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

fc_corona_sel <- fc_corona %>% 
  filter(Country %in% countries_to_forecast &
           Date <= last_date + 8) %>% 
  mutate(Case_Type = case_when(
    Case_Type == "Confirmed" ~ "FC_Confirmed",
    Case_Type == "Deaths" ~ "Exp_growth"))
fc_corona_sel %<>% as_tibble(fc_corona_sel) %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

corona_country_sel <- bind_rows(corona_country_sel, fc_corona_sel)
```

```{r Daily Confirmed lagged, fig.width = 5, fig.asp = 1.2}

data <- corona_country %>%  filter(Country %in% countries_to_forecast &
           Date >= last_date - 21)

p <- ggts_conf_deaths_facet(data, y = Daily_Cases, col = Country)
p # ggplotly(p)
```



```{r }

# identify "best" lag_day with lowest AICc 

lag_days <-  max(countries_fc_param$confirmed_to_death)
lag_range <- lag_days + 21
country <- countries_to_forecast[1]


corona_tsbl_lag_range <- corona_tsbl %>% ungroup() %>% group_by(Case_Type) %>% 
  filter(Country %in% countries_to_forecast &
           Case_Type != "Recovered" &
           Date >= last_date - lag_range) %>% 
  dplyr::select(Date, Country, Case_Type, Daily_Cases) %>% 
  group_by(Country)

```


```{r plot facet daily confirmed and daily deaths, eval = FALSE}

corona_tsbl_lag_range %>% 
  filter(Country == country) %>% 
  ggplot(aes(x = Date, y = Daily_Cases)) +
  geom_line() +
  facet_grid(vars(Case_Type), scales = "free_y") +
  labs(x = "Day", y = NULL,
       title = paste(country, "- Daily Confirmed and Death Cases"),
       subtitle = "w/o lagging and w/o fatality factor")

```

```{r}
## wide format to get Daily Deaths as function of Daily Confirmed Cases
corona_tsbl_lag_range_wide <- corona_tsbl_lag_range %>%   
  pivot_wider(id_cols = c(Country, Date), 
              names_from = Case_Type, 
              values_from = Daily_Cases) %>% 
  tsibble::as_tsibble(index = Date, key = c(Country))
  
no_values <- nrow(corona_tsbl_lag_range_wide %>% filter(Country == country)) 

```


```{r caluclate lag confirmed country dependent}

lm_term <- c("lag(Confirmed, lag_days - 2)",
             "lag(Confirmed, lag_days - 1)",
             "lag(Confirmed, lag_days)",
             "lag(Confirmed, lag_days + 1)",
             "lag(Confirmed, lag_days + 2)")

country_table <- tibble()
save_plot_fit <- list()

## issue testing ##
for (country in countries_to_forecast) {  
  lag_days <- (countries_fc_param %>% 
                 filter(Country == country))$confirmed_to_death
  # model ~ 0 + => w/o intercept
  fit_best <- corona_tsbl_lag_range_wide %>% 
    filter(Country == country) %>% 
    model(ARIMA(Deaths ~ 0 + pdq(d = 0) + 
                  lag(Confirmed, lag_days - 2) +
                  lag(Confirmed, lag_days - 1) +
                  lag(Confirmed, lag_days) +
                  lag(Confirmed, lag_days + 1) +
                  lag(Confirmed, lag_days + 2)
    ))
  if (country == "Germany") {
    fit_best_germany <- fit_best
  }

  corona_deaths_of_confirmed <- corona_tsbl_lag_range_wide  %>% ungroup() %>%
    new_data(lag_days, keep_all = TRUE)
  
  
  ## plot Daily Deaths plus FC of model(ARIMA(Deaths ~ 0 + lag(Confirmed, lag_days)))
  plot_fit <- fit_best %>%
    forecast(corona_deaths_of_confirmed) %>%
    autoplot(corona_tsbl_lag_range_wide %>%  
               filter(Date >= last_date - 20)) + 
    ylab("Daily Deaths") +
    ggtitle(paste(country, "Forecast Daily Deaths based on Confirmed Cases"), 
            subtitle = paste("ARIMA-Model, lag Confirmed:", lag_days, "days"))
  save_plot_fit[[country]] <- plot_fit
  country_table_new <- c(Country = country, 
                         lag_Confirmed = lag_days, 
                         CFR = as.numeric(
                           round(coef(fit_best) %>%  
                                   filter(term %in% term) %>%
                                   summarise(CFR = sum(estimate)), digits = 2))
  )
  country_table <- bind_rows(country_table, country_table_new)
}



```


### Lag days and Case Fatality Rate (CFR)  {data-height=100}

```{r CFR Table - for compare_countries}

title <- paste("Case Fatality Rate in % (CFR, proportion of deaths from confirmed cases)", "lagged by", lag_n, "days; total and for period of past", range_cfr, "days")
knitr::kable(cfr_table %>% filter(Country %in% countries_to_forecast),
             digits = 2, caption = title)



# knitr::kable(country_table,
#              digits = 2, caption =
#                "Lag days and CFR (proportion of deaths from confirmed cases)")


# knitr::kable(coef(fit_best) %>% 
#                group_by(Country) %>% filter(term %in% lm_term) %>%  
#                summarise(CFR = sum(estimate) ), 
#              digits = 2, caption = paste(
#                "Case Fatality Fate in % (CFR, proportion of deaths from confirmed cases)", "lagged by", lag_n, "total and for period of past ", lag_n, "days")
```



Column {data-width=600}
-------------------------------------

### Daily Deaths depending on lagged Daily Confirmed Cases


```{r plot lag confirmed country dependent, fig.width = 5, fig.asp = 0.4}

save_plot_fit$Germany

save_plot_fit$Italy

save_plot_fit$Spain

```


### Exampla Germany - White Noise of Forecast Residuals  {data-height=150}

Forecast residuals indicate quality of fit with Arima model:


```{r corona ARIMA tsresiduals, fig.width = 5, fig.asp = 0.618}

# `r coef(fit_best)$.model[1]`

fit_best_germany  %>% 
  gg_tsresiduals() +
  labs(title = paste("Germany -  Forecast Residuals w/ Correlogram and Histogram"),
       subtitle = paste("Model: ARIMA w/ lagged predictor; lag_days:", lag_germany))
```

```{r}

```



References {#bib}
=====================================

***
***
  
### Data Source {data-height=400}

**Data Source**

Data files are provided by **Johns Hopkins University** on GitHub  
<https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series>

* Data files:  
    + *time_series_covid19_confirmed_global.csv* 
    + *time_series_covid19_deaths_global*
    + *time_series_covid19_recovered_global.csv*

The data are visualized on their excellent Dashboard  
**Johns Hopkins University Dashboard**  
<https://coronavirus.jhu.edu/map.html>


**Code Source**

Directory with all R Sources is replicated in GitHub repository:  
<https://github.com/WoVollmer/R-TimesSeriesAnalysis/tree/master/Corona-Virus>

Code is based on ideas from <https://rpubs.com/TimoBoll/583802>

### Bibliography

