---
title: "Coronavirus SARS-CoV-2 / Covid-19"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
link-citations: yes
cite-color: "green"
bibliography: References_Corona.bib
output:
  flexdashboard::flex_dashboard:
  vertical_layout: scroll
params:
  device: "desktop"
---

Bar Chart {data-navmenu="Bar Chart"}
=====================================

***
***

```{r bar chart page titles}

title_bc_cumulative <- paste("Bar Chart (descending order) and Data Table")
title_bc_cum_100k <- paste("Bar Chart -",  title_cum_100k)
title_bc_mean_daily <- paste("Bar Chart -",  title_mean_daily)
title_bc_cfr <- paste("Bar Chart -",  title_cfr)
  
```


```{r}

# used by bar chart page only
corona_last_abs_wide <- corona_data_last_wide %>% 
  dplyr::select(!contains(c("_100k", "CFR")), -Population)

corona_last_rel_wide <- corona_data_last_wide %>% 
  dplyr::select(Country, Date, contains("_100k"))

```

**`r title_bc_cumulative`**
  
Column {data-width=300}
-------------------------------------

###

```{r Top Ten Bar Chart Confirmed}

# plot bar chart deaths / country

case_type <- "Confirmed"
abscissa <- "Cases"
title <- paste(case_type , "- cumulative Cases")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Country != "World" & Case_Type == case_type),
  y = abscissa, title = title, n = 15)
# for cumulative cases world = sum of all => filter out

```

###

```{r Top Ten Bar Chart Deaths}
### Bar Chart Deaths - cumulative  Cases

# plot bar chart deaths / country
case_type <-  "Deaths"

title <- paste(case_type , "- cumulative Cases")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Country != "World" & Case_Type == case_type),
  y = abscissa, title = title, n = 15)

```


Column
-------------------------------------

### Confirmed and Deaths - cumulative and daily Cases (absolute date)

Case numbers are taken from [[@JHU2020a]](#bib). In order to compensate for the 
daily fluctuations, the 
mean number of cases over the past seven day (Mean_Daily_*) are added.

```{r print datatable in wide format}

scroll_y <- "700px"
DT::datatable(corona_last_abs_wide, 
              options = list(pageLength = 50, scrollY=scroll_y))

```


Bar Chart / Inhabitants {data-navmenu="Bar Chart"}
=====================================

***
***
  
**`r title_bc_cum_100k`**

Column {data-width=300}
-------------------------------------

###

```{r Top Ten Bar Chart per 100k Confirmed}

# plot bar chart Confirmed / 100k for countries and + world

case_type <- "Confirmed"
abscissa <- "Cases_100k"
title <- paste(case_type , 
               "- cumulative Cases / 100k")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Case_Type == case_type),
  y = abscissa, title = title, n = 15)

```

###

```{r Top Ten Bar Chart per 100k Deaths}

# plot bar chart deaths / country
case_type  <- "Deaths"

title <- paste(case_type , 
               "- cumulative Cases / 100k")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Case_Type == case_type),
  y = abscissa, title = title, n = 15)

```

Column
-------------------------------------

### Confirmed and Deaths - cumulative and daily Cases per 100,000 Inhabitants (relative data)

Population numbers are taken from [[@UNO2020]](#bib).
In order to compensate for the daily fluctuations, the
mean number of cases over the past seven days (Mean_Daily_*) are added.

```{r print datatable 100k in wide format}

DT::datatable(corona_last_rel_wide, 
              options = list(pageLength = 50, scrollY=scroll_y))

```

Bar Chart / Hot Spots {data-navmenu="Bar Chart"}
=====================================

***
***
  
**`r title_bc_mean_daily`**

Column {data-width=300}
-------------------------------------

###

```{r Top Ten Bar Chart Hot Spot Confirmed}

# plot bar chart  Daily Confirmed Mean / 100k for countries and + world
case_type <- "Confirmed"
abscissa <- "Mean_Daily_Cases_100k"
title <- paste(case_type , 
               "- Mean Daily Cases / 100k")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Case_Type == case_type),
  y = abscissa, title = title, n = 15)

```


###

```{r Top Ten Bar Chart Hot Spot Deaths}

# plot bar chart  Daily Deaths Mean / 100k for countries and + world
case_type <- "Deaths"

title <- paste(case_type , 
               "- Mean Daily Cases / 100k")
hc_bar_chart_country(
  dplyr::filter(corona_data_last, Case_Type == case_type),
  y = abscissa, title = title, n = 15)

```

Column
-------------------------------------

### Confirmed and Deaths - cumulative and daily Cases per 100,000 Inhabitants (relative data)

Population numbers are taken from [[@UNO2020]](#bib).
In order to compensate for the daily fluctuations, the 
mean number of cases over the past seven days (Mean_Daily_*) are added.

```{r print mean datatable 100k in wide format}

DT::datatable(corona_last_rel_wide,
              options = list(pageLength = 50, scrollY=scroll_y))

```

Bar Chart / CFR {data-navmenu="Bar Chart" #cfr}
=====================================

***
***
  
**`r title_bc_cfr`**

Column {data-width=300}
-------------------------------------

###

```{r Top n Bar Chart CFR_past_period}

# plot bar chart CFR Confirmed Mean / Deaths Mean
abscissa <- "CFR_past_period"
title <- paste(abscissa, "- Case Fatality Rate (in %)")
hc_bar_chart_country(corona_data_last_wide, y = abscissa, title = title, n = 15)

```


###

```{r Top n Bar Chart CFR_total or selected countries}

# plot bar chart CFR Confirmed Mean / Deaths Mean
abscissa <- "CFR_total"
title <- paste(abscissa, "- Case Fatality Rate (in %)")
hc_bar_chart_country(corona_data_last_wide, y = abscissa, title = title, n = 15)

# plot bar chart Confirmed - CFR countries and + world
# title <- paste(abscissa, "(in %) - selected countries")
# hc_bar_chart_country(dplyr::filter(cfr_table, 
#                          Country %in% c(compare_countries, "United Kingdom")), 
#                          y = abscissa, title = title)

```

Column  
-------------------------------------

### Case Fatality Rate - Proportion of deaths from confirmed cases {data-height=170}

The number of confirmed cases is an early predictor of the number of deaths.
The number of today's deaths is already determined by the infections about by 
$\sim19$ days ago or respectively by the confirmed cases about by $\sim11$ days 
ago [[see @RWI2020]](#bib).

An average duration of *Confirmed Infection to Death* of $12$ (lag-)days is 
assumed (country-independent for the sake of simplicity) for the calculations, 
since many tests are carried out in the meantime before symptoms appear.

However, this varies considerably depending on country-specific test rate and 
health system. In the worst health systems it may be only one day between
infection confirmation and death, 
the "Confirmed" cases must be "lagged" by $\sim1$ day. 
In the best case, the time from the end of incubation period (in average 
$\sim5-6$ days) to death is an average $\sim14$ days. In this case, the average
*Confirmed infection to Death* period is $\sim14$ days), the "Confirmed" cases
must be correctly "lagged" by $\sim14$ days.
For the assumed time periods see [[@RKI2020a]](#bib), [[@RKI2020b]](#bib), 
for Case Fatality Rate and Incubation Period in general 
see [[@Wikipedia2020a]](#bib), [[@Wikipedia2020b]](#bib).

The simple calculation with unlagged cumulative confirmed cases divided by
cumulative deaths results in a significant underestimation of the CFR in health systems with early disease detection. 
If the number of cumulative cases is already large compared to the number of 
active cases (~ cases from the past two weeks), the "lagged" rsp. "unlagged"
values converge.

The Infection Fatality Rate (IFR) is the fatality rate of all infection, that means
detected confirmed cases and undetected cases (asymptomatic and not tested group).
This lethality is assumed to 
be country independent and only rough estimates exist (RKI: bottom of existing estimates $\sim0.56\%$). 


```{r CFR Table}

title <- paste("Case Fatality Rate (in %) - assuming a time lag of", lag_n, 
               "days between Confirmed => Death for",
               "CFR_past_period (period of ", range_cfr, "days) and CFR_total")

```

### `r title` {data-height=430}

```{r}

DT::datatable(corona_data_last_wide %>% 
                dplyr::select(Country, Date, 
                              CFR_past_period, CFR_total, CFR_unlagged),  
              options = list(pageLength = 25, scrollY = "300px"))

```

### Case Fatality Rate (in %) - selected countries {data-height=200}


```{r CFR Table selected countries}

knitr::kable(corona_data_last_wide %>% 
               dplyr::select(Country, Date, 
                             CFR_past_period, CFR_total, CFR_unlagged) %>% 
               dplyr::filter(Country %in% compare_countries),
             digits = 2)

# DT::datatable(dplyr::filter(cfr_table, Country %in% compare_countries),
#               options = list(pageLength = 10, scrollY = "300px"))
```


Data Table - all Countries {data-navmenu="Bar Chart"}
=====================================

**Countries - Table overview**

Column
-------------------------------------

### Confirmed and Deaths - cumulative and mean daily Cases, overall and per 100,000 Inhabitants (absolute and relative data)

Population numbers are taken from [[@UNO2020]](#bib).
In order to compensate for daily fluctuations, the mean number of cases for the 
past seven days (Mean_Daily_*) is used instead of the daily cases.

```{r print datatable all with 100k in wide format}

# DT::datatable(corona_data_last %>% dplyr::filter(Case_Type != "Recovered"),
#               options=list(pageLength=25))

DT::datatable(
  corona_data_last_wide %>% 
    dplyr::filter(Country != "World") %>% 
    mutate("Population /m" = round(Population / 1000000, digits = 3)) %>% 
    dplyr::select(
      Country, `Population /m`, Date, 
      Confirmed, Mean_Daily_Conf, Conf_100k, Mean_Daily_Conf_100k,
      Deaths, Mean_Daily_Deaths, Deaths_100k, Mean_Daily_Deaths_100k), 
  options = list(pageLength = 50, scrollY=scroll_y),
  filter = "bottom")

```

Data Table - EU {data-navmenu="Bar Chart"}
=====================================

**European Union - Table overview**

Column
-------------------------------------

### Confirmed and Deaths - cumulative and mean daily Cases, overall and per 100,000 Inhabitants (absolute and relative data)

Population numbers are taken from [[@UNO2020]](#bib).
In order to compensate for daily fluctuations, the mean number of cases for the 
past seven days (Mean_Daily_*) is used instead of the daily cases.

```{r print datatable EU with 100k in wide format}

# DT::datatable(corona_data_last %>% dplyr::filter(Case_Type != "Recovered"),
#               options=list(pageLength=25))

DT::datatable(
  corona_data_last_wide %>% 
    dplyr::filter(Country %in% countries_eu) %>% 
    mutate("Population /m" = round(Population / 1000000, digits = 3)) %>% 
    dplyr::select(
      Country, `Population /m`, Date, 
      Confirmed, Mean_Daily_Conf, Conf_100k, Mean_Daily_Conf_100k,
      Deaths, Mean_Daily_Deaths, Deaths_100k, Mean_Daily_Deaths_100k),
  options = list(pageLength = 50, scrollY=scroll_y),
  filter = "bottom")

```

