---
title: "Coronavirus SARS-CoV-2 / Covid-19"
author: "Wolfgang Vollmer"
date: '`r Sys.Date()`'
link-citations: yes
cite-color: "green"
bibliography: References_Corona.bib
output:
  flexdashboard::flex_dashboard:
  vertical_layout: scroll
params:
  device: "desktop"
---


Doubling Time / Forecast Accuracy {data-navmenu="Forecast" #doubling}
=====================================

***
***

Column {data-width=500}
-------------------------------------

**Doubling Time and Forecast** 
The forecasted cases for the next 14 days are calculated 'only' from the
linear regression of the logarithmic data and are not considering any effects of 
measures in place. In addition data inaccuracies are not taken into account, 
especially relevant for the confirmed cases. 

Therefore the 14 days forecast is only an indication for the direction of an 
unchecked exponentiell growth.

The Cumulative cases doubling rate is only a good indicator at the beginning of 
the pandemic. 
If the number of confirmed cases from the past two weeks is already small 
compared to the total number of confirmed cases, the number of infectious people 
is also small compared to the total cases.

Therefore the table below provides the doubling time for the 
daily rolling mean cases. The forecast is based on the linear regression of the 
logarithmic data of past `r range_reg_long` days.

```{r daily rolling mean cases forecast coef beta_0 and beta_1, eval = TRUE}

fit_corona <- corona_range_reg_long_tsbl %>%  
  model(TSLM(log(Daily_Cases_Mean) ~ Date)) 

coefs <- tidy(fit_corona) %>% 
  group_by(Country, Case_Type) %>% 
  mutate(T_doubling = (log(2) /estimate))

forecast_table <- coefs %>% 
  pivot_wider(id_cols = c(Country, Case_Type), 
              names_from = term, 
              values_from = estimate) %>% 
  group_by(Country, Case_Type) %>%
  rename(beta_0 = `(Intercept)`,
         beta_1 = Date) %>% 
  mutate(T_doubling = (log(2) / beta_1),
         Reg_last_day = exp(beta_0 + unclass(last_date) * beta_1),
         FC_next_day = exp(beta_0 + unclass(last_date + 1) * beta_1),
         FC_14days = exp(beta_0 + unclass(last_date + 1 + range_forcast-1) 
                         * beta_1)) %>% 
  arrange(Case_Type) %>% 
  select(Country, Case_Type, T_doubling, Reg_last_day, 
         FC_next_day, FC_14days)

knitr::kable(forecast_table,  digits = c(0, 0, 1, 0, 0),
             format.args = list(big.mark = "'"),
             caption = 
             "Forecast (FC) Daily Rolling Mean Cases: Doubling Time (days), 
            Forecasted cases next day/tomorrow and  Forecasted cases in 14 days")

```


Column
-----------------------------------------------------------------------  

### Forecast of Daily  and Cumulative Cases and check of Accuracy {data-height=250}

The forecast accuracy is checked by using the forecast method for the past 
three weeks before the past week (*training data*). Subsequent forecasting of the 
past week enables comparison with the real data of these days (*test data*).

The comparison is also an early indicator if the *exponential growth* is 
declining. However, possible changes in underreporting
(in particular the proportion confirmed / actually infected) requires careful 
interpretation.

For doubling periods of the total cumulative cases in the order of infectivity (RKI assumption: $\sim9-10$
days, with great uncertainty, [[see @RKI2020b]](#bib), we have no  
exponential growth for the total cumulative cases. Since the "old" infected cases are 
no longer infectious after these periods and we then have a constant infection 
rate with  basic reproduction number $R_t \sim 1$.

Instead, we have "only" linear growth of the cumulative *Confirmed Cases* and 
the *Daily Confirmed Cases* remain more or less constant if $R_t \sim 1$.

However, the basic reproduction number 
$R_0 (\approx 3.3 - 3.8)$ [[@RKI2020c]](#bib) is a product of the average number
of contacts of an infectious person per day, the probability of transmission upon contacts and the average number of days infected people are infectious. 
With the current uncertainty of the average duration of the infectivity duration, $R_0$ can therefore be estimated from the doubling time only to a very limited extent.
See also [[@CMMID2020]](#bib).

### Germany - Forecast Daily Rolling Mean Cases based on past `r range_reg_long` days {data-height=225}


```{r forecast daily exp model, fig.asp = fig_asp_trend}

fc_daily_corona %>%  
  dplyr::filter(Country == country_select & .model == "Exp_Growth") %>% 
  autoplot(corona_range_reg_long_tsbl) +
  labs(title = paste(
    country_select, " - Forecast of Mean Daily Cases for the next ", 
    forecast_range, "-days"),
    subtitle = paste(
      "Forecast is based on regression of past ",
      range_reg_long, " days")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left")

```

```{r forecast check settings}
# Note: as long as zero deaths are in the training data time range 
# log(0) will go to infinity => no linear regression possible

# total = training + test_data => 
#    Date >= last_date + 1 - training_data_start = last training_data_start days

training_data_start <- range_reg_long +  forecast_range 
              # (last_date + 1) minus x days
forecast_days <- paste(forecast_range, "days")

corona_data_train <- corona_data %>% 
  group_by(Case_Type) %>% # since .add = FALSE will override existing groups
  dplyr::filter(Country %in% country_select & 
           Date >= last_date + 1 - training_data_start &
           Date < last_date + 1 - forecast_range)

corona_data_total <- corona_data %>% 
  group_by(Case_Type) %>% 
  dplyr::filter(Country %in% country_select  & 
           Date >= last_date + 1 -  training_data_start) 

```

### Germany - Forecast Accuracy Check Daily Rolling Mean Cases for the past `r forecast_range` days {data-height=225}

```{r daily forecast accuracy, fig.asp = fig_asp_trend}

fit_corona_train <- corona_data_train %>% 
  model(TSLM(log(Daily_Cases_Mean) ~ Date)) 
fc_corona_train <- fit_corona_train %>% 
  fabletools::forecast(h = forecast_days) 

# forecasts accuracy
# for evaluating accuracy on forecasts to be provided:
# - a complete dataset that includes the future data and 
# - data used to train the model.

# level = prediction interval in % or NULL
fc_corona_train %>%
  autoplot(corona_data_total) +
  labs(title = paste(
    country_select, "- Forecast based on regression of previous",
    range_reg_long, "days"),
    subtitle = "Forecast (blue line) vs. Data (black line)") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left") +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days")

```


```{r cumulative forecast accuracy, fig.asp = fig_asp_trend, eval = FALSE}
### Germany - Forecast Accuracy Check for the past `r forecast_range` days 
### for the cumulative Cases {data-height=225}

fit_corona_train <- corona_data_train %>% 
  model(TSLM(log(Cases) ~ Date)) 
fc_corona_train <- fit_corona_train %>% 
  fabletools::forecast(h = forecast_days) 

# forecasts accuracy
# for evaluating accuracy on forecasts to be provided:
# - a complete dataset that includes the future data and 
# - data used to train the model.

# level = prediction interval in % or NULL
fc_corona_train %>%
  autoplot(corona_data_total) +
  labs(       title = paste(
    country_select, "- Forecast based on regression of previous",
    range_reg_long, "days"),
    subtitle = "Forecast (blue line) vs. Data (black line)") +
  facet_wrap(~ Case_Type, ncol = 2,  scales = "free_y",
             strip.position = "left") +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 days")

```


Forecast w/ CFR {data-navmenu="Forecast"}
=====================================

***
***

Column {data-width=400}
-------------------------------------

### Forecasting with lagged Predictors

The number of confirmed cases is an early predictor of the number of deaths.
The number of today's deaths is already determined by the infections about by 
$\sim19$ days ago or respectively by the confirmed cases about by $\sim11$ days
ago [see Bar Chart - CFR](#cfr).

The country specific case fatality rate (CFR, proportion of deaths from 
confirmed cases) indicates the country specific testing rate and may depend on quality/capacity of hospitals. 

Overall a rough conclusion on the country specific **underreporting rate** 
(lack of diagnostic confirmation; proportion of all infected to confirmed cases)
is feasible if the infection fatality rate (IFR, confiremd cases plus all 
asymptomatic and undiagnosed infections) is assumed to be country independent 
and the IFR is known (bottom of existing estimates $\sim0.56\%$, assumption by 
RKI see [[@RKI2020b]](#bib).

In this case an estimation of the CFR of $0.06$ $(6\%)$ indicates an
underreporting by a by a factor of $\sim10$. 
A CFR of $0.20$ $(20\%)$ indicates an underreporting by a by a factor of $\sim30$.
This corresponds to RKI assumption of a underreporting by a factor of $11-20$ [[@RKI2020a]](#bib). Unfortunately, the IFR or lethality is still far too 
imprecise for concrete conlusions.


In the model paper RKI assumes for the

* Incubation period $\sim5-6$ days - Day of infection day until symptoms are upcoming)
* Hospitalisation $+4$ days  - Admission to the hospital (if needed) after Incubation Period)
* Average period to death $+11$ - if the patient dies, it takes an average of 
$11$ days after admission to the hospital

Depending on the country-specific test frequency (late or early tests), 
the 

*lag_days - time from receipt of the confirmed test result to death, Confirmed to Death, 
is about $11-13$ days.


Note: these methods are also used for example for advertising campaigns. The
campaign impact on sales will be some time beyond the end of the campaign, and 
sales in one month will depend on the advertising expenditure in each of the 
past few months [[see @Hyndman2020]](#bib).


Column {data-width=600}
-------------------------------------

### Daily Confirmed and Death Cases

```{r forecast for selected countries and lags}

# infected_to_confirmed  # infected -> confirmed ~ 9 days,
#                        # Incubation period ~ 5-6 days + testing + get result
#                        # Hospitalisation + 4 days
# infected_to_death      # infected -> death     ~ after 21 days
# lag_x : confirmed_to_death     # ~ 13 days, ~ 21 - 9 days
# case_fatality_rate     # proportion of deaths / number of people diagnosed
# 
#                                                                   (Confirmed)

lag_germany <- 12 # 13
lag_italy <-  12  # 10
lag_spain <- 12   # 11

countries_to_forecast <- c("Germany", "Italy", "Spain")

countries_fc_param <-
  tibble(Country = countries_to_forecast,
       confirmed_to_death = c(lag_germany, lag_italy , lag_spain),
       infected_to_death     = 21) %>% 
       # case_fatality_rate = c(0.08, 0.20, 0.13))
  mutate(infected_to_confirmed = infected_to_death  - confirmed_to_death)

        
```


```{r eval = TRUE}

corona_country_sel <- corona_data %>% 
  dplyr::filter(Country %in% countries_to_forecast & 
           Date >= last_date - 28) %>%
  ungroup() %>% 
  dplyr::select(Date, Country, Case_Type, Cases)

fc_corona_sel <- fc_corona %>% 
  dplyr::filter(Country %in% countries_to_forecast &
           Date <= last_date + 8) %>% 
  mutate(Case_Type = case_when(
    Case_Type == "Confirmed" ~ "FC_Confirmed",
    Case_Type == "Deaths" ~ "Exp_growth"))
fc_corona_sel %<>% as_tibble(fc_corona_sel) %>% 
  dplyr::select(Date, Country, Case_Type, Daily_Cases_Mean)

corona_country_sel <- bind_rows(corona_country_sel, fc_corona_sel)
```

```{r Daily Confirmed lagged, fig.width = 5, fig.asp = 1.2}

data <- corona_data %>%  dplyr::filter(Country %in% countries_to_forecast &
           Date >= last_date - 21)


p <- ggts_conf_deaths_facet(data, y = Daily_Cases_Mean) +
      ggtitle("Daily Rolling Mean Cases (past 3 weeks)")
p # ggplotly(p)
```



```{r }

# identify "best" lag_day with lowest AICc 

lag_days <-  max(countries_fc_param$confirmed_to_death)
lag_range <- lag_days + 21
country <- countries_to_forecast[1]


corona_data_lag_range <- corona_data %>% ungroup() %>% group_by(Case_Type) %>% 
  dplyr::filter(Country %in% countries_to_forecast &
           Date >= last_date - lag_range) %>% 
  dplyr::select(Date, Country, Case_Type, Daily_Cases_Mean) %>% 
  group_by(Country)

```


```{r}
## wide format to get Daily Deaths as function of Daily Confirmed Cases
corona_data_lag_range_wide <- corona_data_lag_range %>%   
  pivot_wider(id_cols = c(Country, Date), 
              names_from = Case_Type, 
              values_from = Daily_Cases_Mean)

```


```{r caluclate lag confirmed country dependent}

lm_term <- c("lag(Confirmed, lag_days - 2)",
             "lag(Confirmed, lag_days - 1)",
             "lag(Confirmed, lag_days)",
             "lag(Confirmed, lag_days + 1)",
             "lag(Confirmed, lag_days + 2)")

country_table <- tibble()
save_plot_fit <- list()

## issue testing ##
for (country in countries_to_forecast) {  
  lag_days <- (countries_fc_param %>% 
                 dplyr::filter(Country == country))$confirmed_to_death
  # model ~ 0 + => w/o intercept
  fit_best <- corona_data_lag_range_wide %>% 
    dplyr::filter(Country == country) %>% 
    model(ARIMA(Deaths ~ 0 + pdq(d = 0) + 
                  lag(Confirmed, lag_days - 2) +
                  lag(Confirmed, lag_days - 1) +
                  lag(Confirmed, lag_days) +
                  lag(Confirmed, lag_days + 1) +
                  lag(Confirmed, lag_days + 2)
    ))
  if (country == "Germany") {
    fit_best_germany <- fit_best
  }

  corona_deaths_of_confirmed <- corona_data_lag_range_wide  %>% 
    ungroup() %>%
    new_data(lag_days, keep_all = TRUE)
  
  
  ## plot Daily Deaths plus FC of model(ARIMA(Deaths ~ 0 + lag(Confirmed, lag_days)))
  plot_fit <- fit_best %>%
    forecast(corona_deaths_of_confirmed) %>%
    #  forecast horizon automatically (by new data) fixed to lag-days minus 
    #     model dependency lag_days - 1 & lag_days - 2 => 12days - 2 = 10 days
    autoplot(corona_data_lag_range_wide %>%  
               dplyr::filter(Date >= last_date - 20)) + 
    labs(y = "Daily Deaths",
         subtitle = 
           paste(country, "w/ avg. Confirmed -> Death:", lag_days, "lag-days"))
  save_plot_fit[[country]] <- plot_fit
  country_table_new <- c(Country = country, 
                         lag_Confirmed = lag_days, 
                         CFR = as.numeric(
                           round(coef(fit_best) %>%  
                                   dplyr::filter(term %in% term) %>%
                                   summarise(CFR = sum(estimate)), digits = 2))
  )
  country_table <- bind_rows(country_table, country_table_new)
}

plot_deaths_forecast <- 
  save_plot_fit$Germany / save_plot_fit$Italy / save_plot_fit$Spain + 
  plot_annotation(
    title = "Forecast Mean Daily Deaths for next 10 days") +
  plot_layout(guides = 'collect') &
  theme(legend.position='bottom')
plot_deaths_forecast 

```


### Lag days and Case Fatality Rate (CFR)  {data-height=100}

```{r CFR Table - for compare_countries}

title <- paste("Case Fatality Rate in % (CFR, proportion of deaths from confirmed cases)", "lagged by", lag_n, "days; total and for period of past", range_cfr, "days")
knitr::kable(cfr_table %>% dplyr::filter(Country %in% countries_to_forecast),
             digits = 2, caption = title)



# knitr::kable(country_table,
#              digits = 2, caption =
#                "Lag days and CFR (proportion of deaths from confirmed cases)")


# knitr::kable(coef(fit_best) %>% 
#                group_by(Country) %>% dplyr::filter(term %in% lm_term) %>%  
#                summarise(CFR = sum(estimate) ), 
#              digits = 2, caption = paste(
#                "Case Fatality Fate in % (CFR, proportion of deaths from confirmed cases)", "lagged by", lag_n, "total and for period of past ", lag_n, "days")
```



Column {data-width=600}
-------------------------------------

### Daily Deaths Forecast depending on lagged Daily Confirmed Cases (ARIMA model)


```{r plot lag confirmed country dependent, fig.width = 5, fig.asp = 1.2}

plot_deaths_forecast

```


### Exampla Germany - White Noise of Forecast Residuals {data-height=150}

Forecast residuals  
indicate quality of  
Arima model fit:


```{r corona ARIMA tsresiduals, fig.width = 5, fig.asp = 0.618}

# `r coef(fit_best)$.model[1]`

fit_best_germany  %>% 
  gg_tsresiduals() +
  labs(title = paste("Germany -  Forecast Residuals w/ Correlogram and Histogram"),
       subtitle = paste("Model: ARIMA w/ lagged predictor; lag_days:", lag_germany))
```

```{r}

```


